### 2013-10-23 Paper_Draft.tex
Putting data in an accessable location

% This is a sample document using the University of Minnesota, Morris, Computer Science
% Senior Seminar modification of the ACM sig-alternate style. Much of this content is taken
% directly from the ACM sample document illustrating the use of the sig-alternate class. Certain
% parts that we never use have been removed to simplify the example, and a few additional
% components have been added.

% See https://github.com/UMM-CSci/Senior_seminar_templates for more info and to make
% suggestions and corrections.

\documentclass{sig-alternate}
\usepackage{color}

%%%% User-defined macros
\newcommand{\lam}{\lambda}
\newcommand{\mycomment}[1]{\textcolor{red}{#1}}
%%%%% Uncomment the following line and comment out the previous one
%%%%% to remove all comments
%%%%% NOTE: comments still occupy a line even if invisible;
%%%%% Don't write them as a separate paragraph
%\newcommand{\mycomment}[1]{}

\begin{document}

% --- Author Metadata here ---
%%% REMEMBER TO CHANGE THE SEMESTER AND YEAR
\conferenceinfo{UMM CSci Senior Seminar Conference, December 2013}{Morris, MN}

\title{Agile Testing Four Page Draft}

\numberofauthors{1}

\author{
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
\alignauthor
Chris M. Thomas\\
	\affaddr{Division of Science and Mathematics}\\
	\affaddr{University of Minnesota, Morris}\\
	\affaddr{Morris, Minnesota, USA 56267}\\
	\email{thom3706@morris.umn.edu}
}

\maketitle
\begin{abstract}
\mycomment{Needs to be written!}

\mycomment{passages or ideas worth putting into my abstract below}

When it comes to software development perhaps one of the most important and time consuming processes is that of software testing.  In fact, some early studies on software testing estimated that software testing could consume fifty percent or more of development costs for a product.

Currently there is much debate in the testing world about whether or not test first testing or test last testing is superior and what advantages one type of testing may offer over the other.  The goal of this paper is to therefore attempt to resolve this debate as much as possible by analyzing research data in the field concerning the advantages and disadvantages of each of these testing methods.

Due to these issues new methods are starting to appear in the agile community that contain some of TDD's main tenets but have shifted enough away from TDD at some fundamental level that they are starting to receive their own names and classifications. 
\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Delphi theory}

\keywords{ACM proceedings, \LaTeX, text tagging}

\section{Introduction}
When it comes to software development perhaps one of the most important and time consuming processes is that of software testing.  In fact, some early studies on software testing estimated that software testing could consume fifty percent or more of development costs for a product ~\cite{Bertolino;2007}.  because of this business's over time have become increasingly interested in attempting to optimize testing to reduce development costs.

During the 1990's \mycomment{note this is a guess, double check this!} businesses focused on using heavily regimented coding practices that was very process oriented.  Common code methods therefore tended to follow what was called a waterfall model where products were produced by running through a series of phases (IE first there was the planning phase, then designing phase
occurred, and then once the product was designed the testing phase would occur to make sure the product was customer ready. This type of design often encouraged a test last methodology which predominated this era.

In the early 2000's though critics of the heavily regimented coding practices produced a new set coding practices that focused not on heavily regimented coding practices but on more spontaneous and reactive coding practices.  These new coding practices are referred to as "agile" coding practices after the agile manifesto written in (need date here) which highlights many of the main tenets of this new style of coding.  In this new style of coding test last methodologies became less useful because agile methodologies believed that testing should occur while the code was being written, not after.  This gave rise to a new testing strategy called test first testing or Test Driven Development (TDD).

Currently there is much debate in the testing world about whether or not test first testing or test last testing is superior and what advantages one type of testing may offer over the other.  The goal of this paper is to therefore attempt to resolve this debate as much as possible by analyzing research data in the field concerning the advantages and disadvantages of each of these testing methods.  Also because research often lags behind current implementations and the field of agile development is currently changing due to its relatively new implementation this paper also plans to look into where new agile ideas are going and compare that to the research given in order to give the best possible representation of the two differing fields.  This paper will therefore discuss the advantages and disadvantages of TDD testing versus waterfall testing and explore new emerging agile methodologies in an attempt to determine the usefulness of each methodology.

In order to accomplish this we have divided the rest of the paper into four additional sections which will take about the specific ideas above in more detail. In section two of the paper we will discuss what software testing is in the context of this paper and two popular ways of implementing it, test first and test last testing.  In section three we will provide an analysis of the data explaining the potential advantages and disadvantages of TDD compared to Waterfall testing.  In section four we will go over new TDD based testing methodologies and in section five we will state our conclusions.

\section{Background}
\subsection{Software Testing}
Software testing, simply defined, is a branch of software engineering that's main purpose is to either identify potential malfunctions or to demonstrate functionality in a software system by observing the software systems execution ~\cite{Bertolino:2007}.  A simple example of this is running your code after you write it to see if it works. \mycomment{this example needs work}  The main purpose of software testing is that of increasing code quality of software products and is used in tandem with software development techniques to produce serviceable code. Although the purpose of software testing is fairly straight forward (to improve code performance) the understanding on how to best achieve this goal for any given situation can be extraordinarily complex ~\cite{Bertolino:2007}. \mycomment{this transition needs some serious help}  Due to this complexity various methodologies and beliefs currently exist on how to best test products.  In this paper our goal will be to compare various testing methodologies.

\mycomment{this transition between these two paragraphs is rather rocky}

It is important to note that there is no one standard quantitative measurement for code quality because it is a qualitative measurement.  Because of this many different types of measurement are used to attempt to prove that one testing methodology is better then another.  Although there are many different attributes that applicable for arguing the superiority of one methodology over another this paper will only focus on four attributes that are commonly found throughout research; code coverage, time efficiency, customer satisfaction, and code correctness. \mycomment{I'm not sure this paragraphs approach is the best way to get the information I want across to the reader, I probably should probably put more detail into each of my four attributes}

\mycomment{may need to add a small bit about functional testing vs unit testing for section 4}

\subsection{Testing Methods}
In this paper we will be mainly focusing on two different testing domains, waterfall/test last testing and TDD or test first a popular agile testing method.

Waterfall or test last testing is currently a very popular testing method and is usually the first way of testing that people implement.  This method follows the basic principle of a person first writes their code and then proceeds after the code is considered initially implemented to write tests that check the functionality of their code.  The developer then runs the test against the code and changes the code until the code passes the developers tests.  This testing is often used with plan based development or testing on the fly because it allows the developer to only have to write tests about parts of the code he or she is worried about. \mycomment{This paragraph needs some serious clean up and much more detail on the topic but for now it gets the basic point across}

TDD or Test first testing is another popular testing method and is primarily used in the agile community.  This follows the basic principles that one should actually write tests before code and tailor the code to the tests instead of tailoring the tests to the code.  The idea behind this testing style is that by doing this one can write their codes objectives in tests so that your code becomes tailored to your objectives instead of objectives being tailored to your code.  \mycomment{again this paragraph needs some serious clean up and detail but for now it gets the basic point across}

\mycomment{things to add to this section}

-Add a brief history to each section

-Add justification for why TDD is agile or why TDD is used as an agile practice

-Use a more specific TDD example by using extreme programming

-write a brief conclusion paragraph containing:

-That TDD and waterfall testing are vague terms and not specific

-Mention that there will also be a focus Evolved TDD forms briefly(BDD, ATDD)

\section{Analysis of TDD}
\subsection{TDD Stances and Data}
Currently, TDD is the most popular testing method in the Agile community.  Many supporters of TDD claim that it reduces overall time spent on a product, improves test coverage, and increases overall code quality.  More importantly there exists data to back up some of these claims.  for example, in the experiment documented by ~\cite{Lemos:2012}, it was noted that on average TDD methodologies increased code coverage by 40\% compared to conventional testing methods.  Another study, documented in ~\cite{Hellman:2012} also suggested some positive traits about TDD by claiming some research showed that TDD practices could reduce effort by up to 27\%. Studies in articles ~\cite{Hellman:2012, Hammond:2012, Kettunen:2010} acknowledged multiple studies that empirically showed at least a marginal increase in product quality when TDD was implemented.  By looking at this data alone it would be relatively easy to see why TDD has such a positive image in industry and academia.

That being said, TDD also has many detractors as well who would tell you that TDD actually increases time spent on a product while not increasing overall code quality. This side also has a fair bit of research backing its opinion.  For example, in ~\cite{Lemos:2012} the same study that showed that TDD increased test coverage by 40\%, it was noted that TDD took significantly more time to implement then conventional testing and did not actually increase code quality. Also, the same study that found research showing the TDD could reduce effort by up to 27\% found other research that suggested that TDD actually increased the effort by two fold ~\cite{Hellman:2012}. It is also worth noting that articles ~\cite{Hammond:2012, Kettunen:2010} also have data and reports arguing that TDD takes more time then conventional testing.  What is perhaps troubling about the data given here is that most of it directly contradicts the data given in the first paragraph.  This suggests an interesting paradox where even though data given on TDD is statistically significant it is decidedly inconclusive.

\subsection{Analysis of Conflicting Data}
Even though the data surrounding TDD is inconclusive as a whole it should still be noted that information can still be extracted from it.  This is especially true because many of the results were statically significant meaning that some sort of meaningful data was obtained.   
Because most of the data was statistically sound it may be worth considering not that the data was faulty but perhaps that the studies contained potentially key differences from one another.  Thus it may be useful to us to try to understand why the data may be so conflicting as it may lead to useful insights about TDD indirectly.

One of the largest issues for summarizing the effectiveness of TDD is the field itself is very broad.  It is important to note that while I specifically defined the original ideas and methodologies of the original idea of TDD, TDD itself has become an umbrella term that has been used to describe a massive array of different testing practices.  This diversity may help explain why we are getting differing results because the TDD practice in one study may be extraordinary different from the TDD practice in another study.  It is worth noting that in ~\cite{Hammond:2012} a major issue with many TDD research papers was the absence of how the TDD process was implemented. This is a major issue in studying the field of TDD because it is hard to tell if people used similar or different methods of TDD to obtain their results.

Another issue for understanding the effectiveness of TDD that many TDD researchers have run into is that TDD is actually a difficult process to implement correctly.  There were many testimonies throughout various articles ~\cite{Hammond:2012, Hellman:2012, Kettunen:2010} of industry professionals finding the TDD process to be much more difficult to implement correctly compared to traditional testing frameworks.  Also, many articles on the analysis of TDD at some point brings up the topic of the complexity of TDD. This acknowledgement of complexity is important as it may suggest a result bias in TDD experiments based on the skill level of the participants as less experienced participants may be significantly less likely to implement TDD in an efficient or successful manner while more experienced participants may have more success.  In one study at a college, although not significantly analyzed due to a lack of participants, it was noted that Alumni who used TDD produced overall better quality code compared to their conventional testing Alumni counterparts while current students in the college showed almost no difference in code quality between TDD and conventional test methods ~\cite{Lemos:2012}.  This potential bias would wreck havoc on TDD data conclusiveness as a whole as the two major participant types for TDD experiments are industry professionals and college students who have vast gaps in experience difference.

Overall it is very hard to draw conclusions on the effectiveness of TDD. Out of all the data out there there is really only one data point that seems to be incontrovertible which is the fact that TDD always seems to increase overall code coverage.  Perhaps another conclusion that can be made is that TDD seems to produce no worse code quality then traditional waterfall testing.  Also there are some implications from potential confounding factors that not all TDD methods are created equal and experience may play a large role in TDD effectiveness. Everything else on the effectiveness of TDD seems to be inconclusive.
\section{evolutions to TDD}
\mycomment{I feel that this is a much needed and very interesting section for exploring Agile testing but I am becoming increasingly concerned about how well this section may be able to tie into the rest of the paper. I would like to use this section as a useful resource for seeing where TDD is going/ what problems TDD developers are trying to address/ see if any of these new spin offs may solve problems that are coming up in research listed in section 3. Any input on how to tie this section into the paper or how to better reach the goals given above would be useful.}

As can been seen in the above sections there are some issues in the current implementation of TDD testing.  Due to these issues new methods are starting to appear in the agile community that contain some of TDD's main tenets but have shifted enough away from TDD at some fundamental level that they are starting to receive their own names and classifications.  Although there are many of these spin-offs in TDD for this paper we chose to focus on three more popular spin offs, acceptance test driven development, Behavior driven development, and Agile Specification Driven Development.

\subsection{acceptance test driven development}

Perhaps the closest spin off to TDD, acceptance test driven development (ATDD), follows many of the ideas and expectations that regular TDD does, like writing tests first and following and using tests to define the development code.  Unlike TDD though ATDD believes that in order to produce higher quality code and better customer satisfaction customers should write or define "acceptance" tests that must be passed before the product is considered finished.  This part of ATDD is noticeably different from the TDD ideology because in the TDD ideology only the developer should writes the tests and the tests should be written from simplest to hardest.~\cite{Hammond:2012}

Although some people claim ATDD is a step in the right direction for TDD developers their are also many people who have complaints about ATDD.  One of the major complaints is that ATDD is an ideal practice at best because customers will not take the time to create useful acceptance tests.~\cite{Hammond:2012} Yet others state that ATDD is very hard for both customers and developers to implement correctly because their is no effective common language between the two factions.(site new IEEE article here)  Because of these varied complaints some of the promoters of ATDD have moved away from it to support a newer similar TDD spin off, Behavior Driven Development or BDD.

\subsection{Behavior Driven Development}
BDD is a new spin off of TDD that focuses on how to correctly implement the fundamental usefulness of TDD.  In BDD the most important thing that it argues is that TDD is unsuccessful because it is vague and it focuses more on "testing the code" then showing the codes behavior.  Thus BDD argues that solving the problem is to stop thinking about code in terms of testing but to understand code in terms of behavior.  Because of this shift in ideas BDD uses different testing tools, such as JBehave for java or Cucumber for Ruby \cite{Soeken:2012}, compared to TDD that breaks code testing down into standard verbal sentences instead of basic testing language.  The argued pros for this type of language usage is that it actually reduces much of the difficulty of tdd by allowing us to define the purpose of our code in more natural terms thus removing much of the confusion about what is important to test in TDD and the other pro is that now it is easier to communicate goals and functionality of the code to non computer science majors (IE Managers, Customers, Sales Reps, etc.) which allows for higher customer satisfaction.
\mycomment{this is just a blurb on BDD and what it is and what perceived advantages it contains, this paragraph will get reworked as it does not fit in the current paper well}

\mycomment{need to do more research here}

\subsection{Agile Specification-Driven Development}
The new testing style Agile Specification-Driven Development is a hybrid of two older testing styles TDD and DbC (design by contract).

Advantage to merging the DbC system with TDD is it gives the designer a better idea about what is important to test \mycomment{and other advantages should also crop up in research on DbC}

\mycomment{need to do research here, specifically it would be useful to do some research on DbC}
\section{conclusion}
In conclusion with the current information available to us there is no obviously superior testing methods between TDD or test first ideologies and waterfall or test last ideologies.  That being said it is worth noting that their are still differences between the two testing methods and that although there is not a clear overall advantage between the two languages certain traits make make one method of testing better then another method of testing in a specific scenario.  For example when dealing with new and relatively inexperienced programmers the test last methodology may be more efficient then the test first methodology.  Another example may be that you believe code coverage is an important necessity for your product, in this case you would probably want to use test first over test last to test your code.  Therefore until research on the subject becomes more clear it would be wise to consider both testing practices as effective and potentially useful for any project you may choose to undertake.

\mycomment{this paragraph is a rough idea of what might be useful conclusions to draw from section three}

For people who are looking to advance the field of TDD it would be potentially very interesting to see how the new TDD spin-offs such as Behavior Driven Development and Agile Specification-Driven Development compare to that of a well documented TDD process like the one suggested for Extreme Programming.  Another thing that could also help advance the field is to do research comparing the effectiveness of waterfall testing and TDD testing between groups with notably different experience levels as some research tentatively points that this may be a relevant variable.

\mycomment{This paragraph is a rough idea on the further research topic}

\mycomment{conclusion brain storm}

-TDD increases code test coverage

-TDD perhaps takes more skill to preform

-TDD may have more effective and less effective practices

-More study is needed.

-Better study methods should be implimented

\bibliographystyle{abbrv}
% sample_paper.bib is the name of the BibTex file containing the
% bibliography entries. Note that you *don't* include the .bib ending here.
\bibliography{Paper_Draft}  
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references

\end{document}

### 2013-10-23 annotated_bibliography.tex
Putting data in an accessable location

--- 2013-10-23 Paper_Draft.tex
@@
 % This is a sample document using the University of Minnesota, Morris, Computer Science
-% Senior Seminar modification of the ACM sig-alternate style. Much of this content is taken
-% directly from the ACM sample document illustrating the use of the sig-alternate class. Certain
-% parts that we never use have been removed to simplify the example, and a few additional
-% components have been added.
+% Senior Seminar modification of the ACM sig-alternate style to generate a simple annotated
+% bibliography. The idea is that this document is fairly short, consisting of a brief description
+% of your sources and how you intend to use them (or not). Most of the ``content'' of the
+% generated document comes from the bibliography file, including the notes field which will
+% provide the annotations.
 
@@
 \documentclass{sig-alternate}
-\usepackage{color}
-
-%%%% User-defined macros
-\newcommand{\lam}{\lambda}
-\newcommand{\mycomment}[1]{\textcolor{red}{#1}}
-%%%%% Uncomment the following line and comment out the previous one
-%%%%% to remove all comments
-%%%%% NOTE: comments still occupy a line even if invisible;
-%%%%% Don't write them as a separate paragraph
-%\newcommand{\mycomment}[1]{}
 
@@
 
-\title{Agile Testing Four Page Draft}
+\title{Agile Testing Section Draft}
 
@@
 \maketitle
-\begin{abstract}
-\mycomment{Needs to be written!}
-
-\mycomment{passages or ideas worth putting into my abstract below}
-
-When it comes to software development perhaps one of the most important and time consuming processes is that of software testing.
-In fact, some early studies on software testing estimated that software testing could consume fifty percent or more of development costs for a product.
-
-Currently there is much debate in the testing world about whether or not test first testing or test last testing is superior and what advantages one type of testing may offer over the other.
-The goal of this paper is to therefore attempt to resolve this debate as much as possible by analyzing research data in the field concerning the advantages and disadvantages of each of these testing methods.
-
-Due to these issues new methods are starting to appear in the agile community that contain some of TDD's main tenets but have shifted enough away from TDD at some fundamental level that they are starting to receive their own names and classifications.
-\end{abstract}
-
-% A category with the (minimum) three required fields
-\category{H.4}{Information Systems Applications}{Miscellaneous}
-%A category including the fourth, optional field follows...
-\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
-
-\terms{Delphi theory}
 
@@
 
+\section{Senior Seminar Paper Goals}
+The goal of this paper is to explore what entails agile testing and its apparent effectiveness in the practical coding world.
+Specifically I would like to explore the current implementations of agile testing and see how it has evolved overtime, and how agile testing compares to conventional testing.
+My articles: ~\cite{George:2003, Schneider:2013, Lemos:2012, Hammond:2012, Soeken:2012, Kettunen:2010, Sawyer:2009, Bertolino:2007, Talby:2006, Hellman:2012}
+
+\section{Key Points}
+1. Testing is very costly and not always effective for industry so optimizing testing practices is very important.
+This was a key point that was brought up in multiple papers, some of which even stated that prior research showed that up to 50\% of all development funds was invested in testing ~\cite{Bertolino:2007, Kettunen:2010}.
+This point is important for understanding the relevance of this paper and this point is also useful for putting testing advantages and disadvantages in a useful context.
+One article in particular ~\cite{Bertolino:2007} really highlights this but many of the other articles touch on it in one way or another.
+
+2. TDD is the main form of Agile testing
+This is acknowledged in almost every article I found on the topic and though few mention other forms of testing the bulk of the interest and detail seems to be focused on TDD, usually in its original form within Extreme Programming.
+TDD is heavily mentioned in the following articles ~\cite{George:2003, Hammond:2012, Hellman:2012, lemos:2012, Kettunen:2010, Talby:2006}
+
+3. TDD has inconclusive results tied to it
+This key point was brought up a lot in papers when explaining TDD itself or the results of TDD experiments.
+for example some papers say TDD makes a product take longer to produce ~\cite{Lemos:2012, Hammond:2012} while others say it decreases the overall development time ~\cite{Kettunen:2010}.
+Other data points that I ran into that had this same indecisive data problem was whether TDD improved code quality, improved customer satisfaction, and/or created less buggy code.
+Perhaps the only agreed on data point is that TDD tends to produce a massive increase in code coverage and test size (no articles disputed this point and many reputed it).
+In particular papers ~\cite{Hammond:2012, Hellman:2012, Kettunen:2010} seem to spend some time attempting to explain this phenomena and all agree that some of the major reasons they give for why this data is all over the place is A.
+TDD is actually hard to implement correctly, B.
+TDD is a very vague term that is not well defined so many different testing methods are done under the same name.
+
+4. New TDD approaches are evolving to help generate goals to promote customer satisfaction and reduce unneeded complexity.
+This key point is brought up by only two articles directly, ~\cite{Soeken:2012 , Hammond:2012} but it was referenced indirectly in other articles.
+Although there is not much backing for what this point is I think it is an important point to bring up as it is showing the future of agile testing.
+
+\section{Outline Begins here}
+
 \section{Introduction}
-When it comes to software development perhaps one of the most important and time consuming processes is that of software testing.
-In fact, some early studies on software testing estimated that software testing could consume fifty percent or more of development costs for a product ~\cite{Bertolino;2007}.
-because of this business's over time have become increasingly interested in attempting to optimize testing to reduce development costs.
-
-During the 1990's \mycomment{note this is a guess, double check this!} businesses focused on using heavily regimented coding practices that was very process oriented.
-Common code methods therefore tended to follow what was called a waterfall model where products were produced by running through a series of phases (IE first there was the planning phase, then designing phase
-occurred, and then once the product was designed the testing phase would occur to make sure the product was customer ready.
-This type of design often encouraged a test last methodology which predominated this era.
-
-In the early 2000's though critics of the heavily regimented coding practices produced a new set coding practices that focused not on heavily regimented coding practices but on more spontaneous and reactive coding practices.
-These new coding practices are referred to as "agile" coding practices after the agile manifesto written in (need date here) which highlights many of the main tenets of this new style of coding.
-In this new style of coding test last methodologies became less useful because agile methodologies believed that testing should occur while the code was being written, not after.
-This gave rise to a new testing strategy called test first testing or Test Driven Development (TDD).
-
-Currently there is much debate in the testing world about whether or not test first testing or test last testing is superior and what advantages one type of testing may offer over the other.
-The goal of this paper is to therefore attempt to resolve this debate as much as possible by analyzing research data in the field concerning the advantages and disadvantages of each of these testing methods.
-Also because research often lags behind current implementations and the field of agile development is currently changing due to its relatively new implementation this paper also plans to look into where new agile ideas are going and compare that to the research given in order to give the best possible representation of the two differing fields.
-This paper will therefore discuss the advantages and disadvantages of TDD testing versus waterfall testing and explore new emerging agile methodologies in an attempt to determine the usefulness of each methodology.
-
-In order to accomplish this we have divided the rest of the paper into four additional sections which will take about the specific ideas above in more detail.
-In section two of the paper we will discuss what software testing is in the context of this paper and two popular ways of implementing it, test first and test last testing.
-In section three we will provide an analysis of the data explaining the potential advantages and disadvantages of TDD compared to Waterfall testing.
-In section four we will go over new TDD based testing methodologies and in section five we will state our conclusions.
+Here I plan to talk about the idea of agile testing and why anyone should care. I also plan to lay out a road map here for the rest of my paper.
 
@@
 \subsection{Software Testing}
-Software testing, simply defined, is a branch of software engineering that's main purpose is to either identify potential malfunctions or to demonstrate functionality in a software system by observing the software systems execution ~\cite{Bertolino:2007}.
-A simple example of this is running your code after you write it to see if it works.
-\mycomment{this example needs work}  The main purpose of software testing is that of increasing code quality of software products and is used in tandem with software development techniques to produce serviceable code.
-Although the purpose of software testing is fairly straight forward (to improve code performance) the understanding on how to best achieve this goal for any given situation can be extraordinarily complex ~\cite{Bertolino:2007}.
-\mycomment{this transition needs some serious help}  Due to this complexity various methodologies and beliefs currently exist on how to best test products.
-In this paper our goal will be to compare various testing methodologies.
+in this portion I will:
+-Give a basic explanation of software testing
+-Give the main goals of software testing
+-Express the importance of software testing
+-Go into current testing methods
 
-\mycomment{this transition between these two paragraphs is rather rocky}
-
-It is important to note that there is no one standard quantitative measurement for code quality because it is a qualitative measurement.
-Because of this many different types of measurement are used to attempt to prove that one testing methodology is better then another.
-Although there are many different attributes that applicable for arguing the superiority of one methodology over another this paper will only focus on four attributes that are commonly found throughout research; code coverage, time efficiency, customer satisfaction, and code correctness.
-\mycomment{I'm not sure this paragraphs approach is the best way to get the information I want across to the reader, I probably should probably put more detail into each of my four attributes}
-
-\mycomment{may need to add a small bit about functional testing vs unit testing for section 4}
-
-\subsection{Testing Methods}
-In this paper we will be mainly focusing on two different testing domains, waterfall/test last testing and TDD or test first a popular agile testing method.
-
-Waterfall or test last testing is currently a very popular testing method and is usually the first way of testing that people implement.
-This method follows the basic principle of a person first writes their code and then proceeds after the code is considered initially implemented to write tests that check the functionality of their code.
-The developer then runs the test against the code and changes the code until the code passes the developers tests.
-This testing is often used with plan based development or testing on the fly because it allows the developer to only have to write tests about parts of the code he or she is worried about.
-\mycomment{This paragraph needs some serious clean up and much more detail on the topic but for now it gets the basic point across}
-
-TDD or Test first testing is another popular testing method and is primarily used in the agile community.
-This follows the basic principles that one should actually write tests before code and tailor the code to the tests instead of tailoring the tests to the code.
-The idea behind this testing style is that by doing this one can write their codes objectives in tests so that your code becomes tailored to your objectives instead of objectives being tailored to your code.
-\mycomment{again this paragraph needs some serious clean up and detail but for now it gets the basic point across}
-
-\mycomment{things to add to this section}
-
--Add a brief history to each section
-
--Add justification for why TDD is agile or why TDD is used as an agile practice
-
--Use a more specific TDD example by using extreme programming
-
--write a brief conclusion paragraph containing:
-
--That TDD and waterfall testing are vague terms and not specific
-
--Mention that there will also be a focus Evolved TDD forms briefly(BDD, ATDD)
+\subsection{Agile Programming}
+in this portion I will:
+-explain Agile Ideology (make sure to list important tenents like customer communication)
+-explain basic TDD
 
@@
 for example, in the experiment documented by ~\cite{Lemos:2012}, it was noted that on average TDD methodologies increased code coverage by 40\% compared to conventional testing methods.
-Another study, documented in ~\cite{Hellman:2012} also suggested some positive traits about TDD by claiming some research showed that TDD practices could reduce effort by up to 27\%.
+Another study, documented in ~\cite{Hellmann:2012} also suggested some positive traits about TDD by claiming some research showed that TDD practices could reduce effort by up to 27\%.
 Studies in articles ~\cite{Hellman:2012, Hammond:2012, Kettunen:2010} acknowledged multiple studies that empirically showed at least a marginal increase in product quality when TDD was implemented.
@@
 For example, in ~\cite{Lemos:2012} the same study that showed that TDD increased test coverage by 40\%, it was noted that TDD took significantly more time to implement then conventional testing and did not actually increase code quality.
-Also, the same study that found research showing the TDD could reduce effort by up to 27\% found other research that suggested that TDD actually increased the effort by two fold ~\cite{Hellman:2012}.
+Also, the same study that found research showing the TDD could reduce effort by up to 27\% (~\cite{Hellman:2012}) found other research that suggested that TDD actually increased the effort by two fold.
 It is also worth noting that articles ~\cite{Hammond:2012, Kettunen:2010} also have data and reports arguing that TDD takes more time then conventional testing.
@@
 \section{evolutions to TDD}
-\mycomment{I feel that this is a much needed and very interesting section for exploring Agile testing but I am becoming increasingly concerned about how well this section may be able to tie into the rest of the paper.
-I would like to use this section as a useful resource for seeing where TDD is going/ what problems TDD developers are trying to address/ see if any of these new spin offs may solve problems that are coming up in research listed in section 3.
-Any input on how to tie this section into the paper or how to better reach the goals given above would be useful.}
+\subsection{BDD}
+Here I plan to explain what BDD is and why it may be better then TDD.
 
-As can been seen in the above sections there are some issues in the current implementation of TDD testing.
-Due to these issues new methods are starting to appear in the agile community that contain some of TDD's main tenets but have shifted enough away from TDD at some fundamental level that they are starting to receive their own names and classifications.
-Although there are many of these spin-offs in TDD for this paper we chose to focus on three more popular spin offs, acceptance test driven development, Behavior driven development, and Agile Specification Driven Development.
+\subsection{ATDD}
+Here I plan to explain what ATDD is and why it may be better then TDD.
+\section{conclusion}
+Here I plan to talk about the main take away points of this paper
 
-\subsection{acceptance test driven development}
+% The following two commands are all you need to
+% produce the bibliography for the citations in your paper.
+\bibliographystyle{abbrv}
+% annotated_bibliography.bib is the name of the BibTex file containing 
+% all the bibliography entries for this example. Note that you *don't* include the .bib ending
+% in the \bibliography command.
+\bibliography{annotated_bibliography}
 
-Perhaps the closest spin off to TDD, acceptance test driven development (ATDD), follows many of the ideas and expectations that regular TDD does, like writing tests first and following and using tests to define the development code.
-Unlike TDD though ATDD believes that in order to produce higher quality code and better customer satisfaction customers should write or define "acceptance" tests that must be passed before the product is considered finished.
-This part of ATDD is noticeably different from the TDD ideology because in the TDD ideology only the developer should writes the tests and the tests should be written from simplest to hardest.~\cite{Hammond:2012}
-
-Although some people claim ATDD is a step in the right direction for TDD developers their are also many people who have complaints about ATDD.
-One of the major complaints is that ATDD is an ideal practice at best because customers will not take the time to create useful acceptance tests.~\cite{Hammond:2012} Yet others state that ATDD is very hard for both customers and developers to implement correctly because their is no effective common language between the two factions.(site new IEEE article here)  Because of these varied complaints some of the promoters of ATDD have moved away from it to support a newer similar TDD spin off, Behavior Driven Development or BDD.
-
-\subsection{Behavior Driven Development}
-BDD is a new spin off of TDD that focuses on how to correctly implement the fundamental usefulness of TDD.
-In BDD the most important thing that it argues is that TDD is unsuccessful because it is vague and it focuses more on "testing the code" then showing the codes behavior.
-Thus BDD argues that solving the problem is to stop thinking about code in terms of testing but to understand code in terms of behavior.
-Because of this shift in ideas BDD uses different testing tools, such as JBehave for java or Cucumber for Ruby \cite{Soeken:2012}, compared to TDD that breaks code testing down into standard verbal sentences instead of basic testing language.
-The argued pros for this type of language usage is that it actually reduces much of the difficulty of tdd by allowing us to define the purpose of our code in more natural terms thus removing much of the confusion about what is important to test in TDD and the other pro is that now it is easier to communicate goals and functionality of the code to non computer science majors (IE Managers, Customers, Sales Reps, etc.) which allows for higher customer satisfaction.
-\mycomment{this is just a blurb on BDD and what it is and what perceived advantages it contains, this paragraph will get reworked as it does not fit in the current paper well}
-
-\mycomment{need to do more research here}
-
-\subsection{Agile Specification-Driven Development}
-The new testing style Agile Specification-Driven Development is a hybrid of two older testing styles TDD and DbC (design by contract).
-
-Advantage to merging the DbC system with TDD is it gives the designer a better idea about what is important to test \mycomment{and other advantages should also crop up in research on DbC}
-
-\mycomment{need to do research here, specifically it would be useful to do some research on DbC}
-\section{conclusion}
-In conclusion with the current information available to us there is no obviously superior testing methods between TDD or test first ideologies and waterfall or test last ideologies.
-That being said it is worth noting that their are still differences between the two testing methods and that although there is not a clear overall advantage between the two languages certain traits make make one method of testing better then another method of testing in a specific scenario.
-For example when dealing with new and relatively inexperienced programmers the test last methodology may be more efficient then the test first methodology.
-Another example may be that you believe code coverage is an important necessity for your product, in this case you would probably want to use test first over test last to test your code.
-Therefore until research on the subject becomes more clear it would be wise to consider both testing practices as effective and potentially useful for any project you may choose to undertake.
-
-\mycomment{this paragraph is a rough idea of what might be useful conclusions to draw from section three}
-
-For people who are looking to advance the field of TDD it would be potentially very interesting to see how the new TDD spin-offs such as Behavior Driven Development and Agile Specification-Driven Development compare to that of a well documented TDD process like the one suggested for Extreme Programming.
-Another thing that could also help advance the field is to do research comparing the effectiveness of waterfall testing and TDD testing between groups with notably different experience levels as some research tentatively points that this may be a relevant variable.
-
-\mycomment{This paragraph is a rough idea on the further research topic}
-
-\mycomment{conclusion brain storm}
-
--TDD increases code test coverage
-
--TDD perhaps takes more skill to preform
-
--TDD may have more effective and less effective practices
-
--More study is needed.
-
--Better study methods should be implimented
-
-\bibliographystyle{abbrv}
-% sample_paper.bib is the name of the BibTex file containing the
-% bibliography entries. Note that you *don't* include the .bib ending here.
-\bibliography{Paper_Draft}  
-% You must have a proper ".bib" file
-%  and remember to run:
-% latex bibtex latex latex
-% to resolve all references
+% You must have a ".bib" file and remember to run:
+%     pdflatex bibtex pdflatex pdflatex
+% in order to see all the citation references correctly.
 

### 2013-10-23 sample_paper.tex
Putting data in an accessable location

% This is a sample document using the University of Minnesota, Morris, Computer Science
% Senior Seminar modification of the ACM sig-alternate style. Much of this content is taken
% directly from the ACM sample document illustrating the use of the sig-alternate class. Certain
% parts that we never use have been removed to simplify the example, and a few additional
% components have been added.

% See https://github.com/UMM-CSci/Senior_seminar_templates for more info and to make
% suggestions and corrections.

\documentclass{sig-alternate}
\usepackage{color}

%%%% User-defined macros
\newcommand{\lam}{\lambda}
\newcommand{\mycomment}[1]{\textcolor{red}{#1}}
%%%%% Uncomment the following line and comment out the previous one
%%%%% to remove all comments
%%%%% NOTE: comments still occupy a line even if invisible;
%%%%% Don't write them as a separate paragraph
%\newcommand{\mycomment}[1]{}

\begin{document}

% --- Author Metadata here ---
%%% REMEMBER TO CHANGE THE SEMESTER AND YEAR
\conferenceinfo{UMM CSci Senior Seminar Conference, December 2013}{Morris, MN}

\title{Agile Testing Section Draft}

\numberofauthors{1}

\author{
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
\alignauthor
Chris M. Thomas\\
	\affaddr{Division of Science and Mathematics}\\
	\affaddr{University of Minnesota, Morris}\\
	\affaddr{Morris, Minnesota, USA 56267}\\
	\email{thom3706@morris.umn.edu}
}

\maketitle
\begin{abstract}
This paper provides a sample of a \LaTeX\ document which conforms,
somewhat loosely, to the formatting guidelines for the University of
Minnesota, Morris, Computer Science Senior Seminar proceedings.
It is based heavily on (and takes material directly from) a similar document 
illustrating the format of the ACM SIG Proceedings, which we have based our
proceedings format on.

The original ACM document tried to include
\begin{quote}
every imaginable sort
of ``bells and whistles", such as a subtitle, footnotes on
title, subtitle and authors, as well as in the text, and
every optional component (e.g. Acknowledgments, Additional
Authors, Appendices), not to mention examples of
equations, theorems, tables and figures.
\end{quote}

We've removed many of the more esoteric tricks here because either
they'll never be used (e.g., multiple authors) or are used \emph{very}
rarely (e.g., appendices). Refer to the original ACM document for more
of those fancy examples. \mycomment{Needs more work}
\end{abstract}

% A category with the (minimum) three required fields
\category{H.4}{Information Systems Applications}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Delphi theory}

\keywords{ACM proceedings, \LaTeX, text tagging}

\section{Introduction}
The \textit{proceedings} are the records of a conference, and conference
editors like ACM seek to give their conference by-products a uniform,
high-quality appearance. We also would like our proceedings to look highly
professional, so we're borrowing heavily from the ACM formatting guidelines.
These include some rigid
requirements for the format of the proceedings documents: there
is a specified format (balanced  double columns), a specified
set of fonts (Arial or Helvetica and Times Roman) in
certain specified sizes (for instance, 9 point for body copy),
a specified live area (18 $\times$ 23.5 cm [7" $\times$ 9.25"]) centered on
the page, specified size of margins (2.54cm [1"] top and
bottom and 1.9cm [.75"] left and right; specified column width
(8.45cm [3.33"]) and gutter size (.083cm [.33"]).

The good news is, with only a handful of manual
settings\footnote{One of these, the {\texttt{\char'134 alignauthor}} command, you have
already used; another, {\texttt{\char'134 balancecolumns}}, will
be used in your very last run of \LaTeX\ to ensure
balanced column heights on the last page.}, the \LaTeX\ document
class file handles all of this for you.

The remainder of this document is concerned with showing, in
the context of an ``actual'' document, the \LaTeX\ commands
specifically available for denoting the structure of a
proceedings paper, rather than with giving rigorous descriptions
or explanations of such commands. Section~\ref{main} introduces the main
examples of formatting.

\section{The {\secit Body} of The Paper}\label{main}

Typically, the body of a paper is organized
into a hierarchical structure, with numbered or unnumbered
headings for sections, subsections, sub-subsections, and even
smaller sections.  The command \texttt{{\char'134}section} that
precedes this paragraph is part of such a
hierarchy.\footnote{This is the second footnote.  It
starts a series of three footnotes that add nothing
informational, but just give an idea of how footnotes work
and look. It is a wordy one, just so you see
how a longish one plays out.} \LaTeX\ handles the numbering
and placement of these headings for you, when you use
the appropriate heading commands around the titles
of the headings.  If you want a sub-subsection or
smaller part to be unnumbered in your output, simply append an
asterisk to the command name.  Examples of both
numbered and unnumbered headings will appear throughout the
balance of this sample document.

Because the entire article is contained in
the \textbf{document} environment, you can indicate the
start of a new paragraph with a blank line in your
input file; that is why this sentence forms a separate paragraph.

\subsection{Type Changes and {\subsecit Special} Characters}

We have already seen several typeface changes in this sample.  You
can indicate italicized words or phrases in your text with
the command \texttt{{\char'134}textit}; emboldening with the
command \texttt{{\char'134}textbf}
and typewriter-style (for instance, for computer code) with
\texttt{{\char'134}texttt}.  But remember, you do not
have to indicate typestyle changes when such changes are
part of the \textit{structural} elements of your
article; for instance, the heading of this subsection will
be in a sans serif\footnote{A third footnote, here.
Let's make this a rather short one to
see how it looks.} typeface, but that is handled by the
document class file. Take care with the use
of\footnote{A fourth, and last, footnote.}
the curly braces in typeface changes; they mark
the beginning and end of
the text that is to be in the different typeface.

You can use whatever symbols, accented characters, or
non-English characters you need anywhere in your document;
you can find a complete list of what is
available in the \textit{\LaTeX\
User's Guide}\cite{Lamport:LaTeX}.

\subsection{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of
the three are discussed in the next sections.

\subsubsection{Inline (In-text) Equations}
A formula that appears in the running text is called an
inline or in-text formula.  It is produced by the
\textbf{math} environment, which can be
invoked with the usual \texttt{{\char'134}begin. . .{\char'134}end}
construction or with the short form \texttt{\$. . .\$}. You
can use any of the symbols and structures,
from $\alpha$ to $\omega$, available in
\LaTeX\cite{Lamport:LaTeX}; this section will simply show a
few examples of in-text equations in context. Notice how
this equation: \begin{math}\lim_{n\rightarrow \infty}x=0\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsubsection{Display Equations}
A numbered display equation -- one set off by vertical space
from the text and centered horizontally -- is produced
by the \textbf{equation} environment. An unnumbered display
equation is produced by the \textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols
and structures available in \LaTeX; this section will just
give a couple of examples of display equations in context.
First, consider the equation, shown as an inline equation above:
%%%%
\begin{equation*}
\lim_{n\rightarrow \infty}x=0
\end{equation*}
%%%%
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}\sum_{i=0}^{\infty} x + 1\end{displaymath}
and follow it with another numbered equation:
\begin{equation}\sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\subsection{Multi-line formulas}
%%%%%% \[ \] denotes math; array environment is needed for multi-line
%%%%%% format, {c} means that the array has one column, centered 
%%%%%  (alternatives are {l} for left-aligned, {r} for right-aligned)
%%%%%  \\ denotes a new line
\[
\begin{array}{c}
n_1 = \sum_{i = 1}^k a_i \\
n_2 = \prod_{i = 1}^k b_i
\end{array}
\]

\subsection{Citations}
Citations to articles \cite{Aaronson:2005,Garey:1979,Brun:2008} listed
in the Bibliography section of your
article will occur throughout the text of your article.
You should use BibTeX to automatically produce this bibliography;
you simply need to insert one of several citation commands with
a key of the item cited in the proper location in
the \texttt{.tex} file \cite{OM:2008}.
The key is a short reference you invent to uniquely
identify each work; in this sample document, the key is
the first author's surname and a
word from the title.  This identifying key is included
with each item in the \texttt{.bib} file for your article.

The details of the construction of the \texttt{.bib} file
are beyond the scope of this sample document, but more
information can be found in the \textit{Author's Guide},
and exhaustive details in the \textit{\LaTeX\ User's
Guide}.

This article shows only the plainest form
of the citation command, using \texttt{{\char'134}cite}.
This is what is stipulated in the SIGS style specifications.
No other citation format is endorsed or supported.

\subsection{Tables}
Because tables cannot be split across pages, the best
placement for them is typically the top of the page
nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and
the table caption.  The contents of the table itself must go
in the \textbf{tabular} environment, to
be aligned properly in rows and columns, with the desired
horizontal and vertical rules.  Again, detailed instructions
on \textbf{tabular} material
is found in the \textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table 1 is included in the input file; compare the
placement of the table here with the table in the printed
dvi output of this document.

\begin{table}
\centering
\caption{Frequency of Special Characters}
\begin{tabular}{|c|c|l|} \hline
Non-English or Math&Frequency&Comments\\ \hline
\O & 1 in 1,000& For Swedish names\\ \hline
$\pi$ & 1 in 5& Common in math\\ \hline
\$ & 4 in 5 & Used in business\\ \hline
$\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
\hline\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of
the page's live area, use the environment
\textbf{table*} to enclose the table's contents and
the table caption.  As with a single-column table, this wide
table will ``float" to a location deemed more desirable.
Immediately following this sentence is the point at which
Table 2 is included in the input file; again, it is
instructive to compare the placement of the
table here with the table in the printed dvi
output of this document.

\begin{table*}
\centering
\caption{Some Typical Commands}
\begin{tabular}{|c|c|l|} \hline
Command&A Number&Comments\\ \hline
\texttt{{\char'134}alignauthor} & 100& Author alignment\\ \hline
\texttt{{\char'134}numberofauthors}& 200& Author enumeration\\ \hline
\texttt{{\char'134}table}& 300 & For tables\\ \hline
\texttt{{\char'134}table*}& 400& For wider tables\\ \hline\end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!

\subsection{Figures}
Like tables, figures cannot be split across pages; the
best placement for them
is typically the top or the bottom of the page nearest
their initial cite.  To ensure this proper ``floating'' placement
of figures, use the environment
\textbf{figure} to enclose the figure and its caption.

This sample document contains examples of %\textbf{.eps}
%and 
a \textbf{.pdf} file to be displayable with \LaTeX.  More
details on each of these is found in the \textit{Author's Guide}.

\begin{figure}
\centering
\psfig{file=sample_graph.pdf,width =3in}
\caption{A sample graph just spanning one column.}
\end{figure}

As was the case with tables, you may want a figure
that spans two columns.  To do this, and still to
ensure proper ``floating'' placement of tables, use the environment
\textbf{figure*} to enclose the figure and its caption.
\begin{figure*}
\centering
\psfig{file=sample_graph.pdf,width =5.3in}
\caption{A sample graph that needs to span two columns of text.}
\end{figure*}
and don't forget to end the environment with
{figure*}, not {figure}!

It's easiest and you tend to get the best quality if your figures vector graphics
in PDF format. You can include other formats such as PNG, but they will usually
not look nearly as professional, especially when printed on high resolution printers.
\emph{Be vary wary of screen captures from other papers. They tend to look pixelated
and amateurish even at high resolutions.}

\subsection{Theorem-like Constructs}
Other common constructs that may occur in your article are
the forms for logical constructs like theorems, axioms,
corollaries and proofs.  There are
two forms, one produced by the
command \texttt{{\char'134}newtheorem} and the
other by the command \texttt{{\char'134}newdef}; perhaps
the clearest and easiest way to distinguish them is
to compare the two in the output of this sample document:

This uses the \textbf{theorem} environment, created by
the\linebreak\texttt{{\char'134}newtheorem} command:
\newtheorem{theorem}{Theorem}
\begin{theorem}
Let $f$ be continuous on $[a,b]$.  If $G$ is
an antiderivative for $f$ on $[a,b]$, then
\begin{displaymath}\int^b_af(t)dt = G(b) - G(a).\end{displaymath}
\end{theorem}

The other uses the \textbf{definition} environment, created
by the \texttt{{\char'134}newdef} command:
\newdef{definition}{Definition}
\begin{definition}
If $z$ is irrational, then by $e^z$ we mean the
unique number which has
logarithm $z$: \begin{displaymath}{\log e^z = z}\end{displaymath}
\end{definition}

Two lists of constructs that use one of these
forms is given in the
\textit{Author's  Guidelines}.

There is one other similar construct environment, which is
already set up
for you; i.e. you must \textit{not} use
a \texttt{{\char'134}newdef} command to
create it: the \textbf{proof} environment.  Here
is a example of its use:
\begin{proof}
Suppose on the contrary there exists a real number $L$ such that
\begin{displaymath}
\lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
\end{displaymath}
Then
\begin{displaymath}
l=\lim_{x\rightarrow c} f(x)
= \lim_{x\rightarrow c}
\left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
= \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
\frac{f(x)}{g(x)} = 0\cdot L = 0,
\end{displaymath}
which contradicts our assumption that $l\neq 0$.
\end{proof}

Complete rules about using these environments and using the
two different creation commands are in the
\textit{Author's Guide}; please consult it for more
detailed instructions.  If you need to use another construct,
not listed therein, which you want to have the same
formatting as the Theorem
or the Definition\cite{salas:calculus} shown above,
use the \texttt{{\char'134}newtheorem} or the
\texttt{{\char'134}newdef} command,
respectively, to create it.

\subsection*{A {\secit Caveat} for the \TeX\ Expert}
Because you have just been given permission to
use the \texttt{{\char'134}newdef} command to create a
new form, you might think you can
use \TeX's \texttt{{\char'134}def} to create a
new command: \textit{Please refrain from doing this!}
Remember that your \LaTeX\ source code is primarily intended
to create camera-ready copy, but may be converted
to other forms -- e.g. HTML. If you inadvertently omit
some or all of the \texttt{{\char'134}def}s recompilation will
be, to say the least, problematic.

\section{Conclusions}
This paragraph will end the body of this sample document.
Remember that you might still have Acknowledgments or
Appendices; brief samples of these
follow.  There is still the Bibliography to deal with; and
we will make a disclaimer about that here: with the exception
of the reference to the \LaTeX\ book, the citations in
this paper are to articles which have nothing to
do with the present subject and are used as
examples only.

\section{Acknowledgments}

This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance and
what have you.

It is common (but by no means necessary) for students to thank
their advisor, and possibly other faculty, friends, and family who provided
useful feedback on the paper as it was being written.

In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.

% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
% sample_paper.bib is the name of the BibTex file containing the
% bibliography entries. Note that you *don't* include the .bib ending here.
\bibliography{sample_paper}  
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references

\end{document}

### 2013-10-27 Paper_Draft.tex
6 Paper draft, working on TDD definition currently in section 2.

@@
 \section{Introduction}
-When it comes to software development perhaps one of the most important and time consuming processes is that of software testing.
-In fact, some early studies on software testing estimated that software testing could consume fifty percent or more of development costs for a product ~\cite{Bertolino;2007}.
-because of this business's over time have become increasingly interested in attempting to optimize testing to reduce development costs.
+When it comes to software development, perhaps one of the most important and time consuming processes is software testing.
+In fact, some early studies on software testing estimated that it could consume fifty percent or more of development costs for a product ~\cite{Bertolino;2007}.
+because of this businesses over time have become increasingly interested in attempting to optimize testing to reduce development costs.
 
-During the 1990's \mycomment{note this is a guess, double check this!} businesses focused on using heavily regimented coding practices that was very process oriented.
-Common code methods therefore tended to follow what was called a waterfall model where products were produced by running through a series of phases (IE first there was the planning phase, then designing phase
-occurred, and then once the product was designed the testing phase would occur to make sure the product was customer ready.
-This type of design often encouraged a test last methodology which predominated this era.
+Although there are many testing methods that currently exist two particular testing methods are currently very popular: test-last testing and test-first testing.
+Test-last testing, used mostly in process oriented development, is a testing method where testing is done after the software code has been written to ensure that the code works correctly.
+Test-first testing, used mostly in interval or sprint oriented development, is a testing method where testing is done before the software code is written to ensure that the future code meets certain requirements.
 
-In the early 2000's though critics of the heavily regimented coding practices produced a new set coding practices that focused not on heavily regimented coding practices but on more spontaneous and reactive coding practices.
-These new coding practices are referred to as "agile" coding practices after the agile manifesto written in (need date here) which highlights many of the main tenets of this new style of coding.
-In this new style of coding test last methodologies became less useful because agile methodologies believed that testing should occur while the code was being written, not after.
-This gave rise to a new testing strategy called test first testing or Test Driven Development (TDD).
+Currently, there is much debate in the testing community about whether or not test first or test last testing is superior.
+The goal of this paper is to attempt to give an overview of the current state of this debate by analyzing current research data concerning the advantages and disadvantages of each of these testing methods.
+Also because research often lags behind current implementations and the field of test first testing is currently changing due to its relatively new implementation this paper will also explore where new test first methods are going and compare them to the research given.
+This paper will therefore discuss the advantages and disadvantages of test first testing versus test last testing and explore new emerging test-first methodologies in an attempt to determine the applicability of each methodology.
 
-Currently there is much debate in the testing world about whether or not test first testing or test last testing is superior and what advantages one type of testing may offer over the other.
-The goal of this paper is to therefore attempt to resolve this debate as much as possible by analyzing research data in the field concerning the advantages and disadvantages of each of these testing methods.
-Also because research often lags behind current implementations and the field of agile development is currently changing due to its relatively new implementation this paper also plans to look into where new agile ideas are going and compare that to the research given in order to give the best possible representation of the two differing fields.
-This paper will therefore discuss the advantages and disadvantages of TDD testing versus waterfall testing and explore new emerging agile methodologies in an attempt to determine the usefulness of each methodology.
-
-In order to accomplish this we have divided the rest of the paper into four additional sections which will take about the specific ideas above in more detail.
-In section two of the paper we will discuss what software testing is in the context of this paper and two popular ways of implementing it, test first and test last testing.
-In section three we will provide an analysis of the data explaining the potential advantages and disadvantages of TDD compared to Waterfall testing.
-In section four we will go over new TDD based testing methodologies and in section five we will state our conclusions.
+The rest of the paper is divided into four sections which will discuss the specific ideas above in more detail.
+In section two of the paper we will discuss what software testing is and current software development models with their supporting testing methods: test-first and test-last testing.
+In section three we will provide an analysis of the data explaining the potential advantages and disadvantages of test first testing compared to test last testing.
+In section four we will go over new test first based testing methodologies and in section five we will state our conclusions and suggestions for further research in the field.
 
@@
 \subsection{Software Testing}
-Software testing, simply defined, is a branch of software engineering that's main purpose is to either identify potential malfunctions or to demonstrate functionality in a software system by observing the software systems execution ~\cite{Bertolino:2007}.
-A simple example of this is running your code after you write it to see if it works.
-\mycomment{this example needs work}  The main purpose of software testing is that of increasing code quality of software products and is used in tandem with software development techniques to produce serviceable code.
-Although the purpose of software testing is fairly straight forward (to improve code performance) the understanding on how to best achieve this goal for any given situation can be extraordinarily complex ~\cite{Bertolino:2007}.
-\mycomment{this transition needs some serious help}  Due to this complexity various methodologies and beliefs currently exist on how to best test products.
-In this paper our goal will be to compare various testing methodologies.
+Software testing, simply defined, is a branch of software engineering that entails a series of practices meant to either identify potential malfunctions or demonstrate functionality in a software system ~\cite{Bertolino:2007}.
+Software testing can be as simple as running a program to see its results or can be as complex as writing code to simulate scenarios in the real world.
 
-\mycomment{this transition between these two paragraphs is rather rocky}
+When it comes to judging the effectiveness of certain testing methods it is important to note that there is no one standard quantitative measurement.
+Because of this many different types of measurement are used to attempt to prove that one testing methodology is better then another.
+Although there are many different attributes that are used for arguing the superiority of one method over another we will only focus on three attributes that are commonly found throughout research; code coverage, time, and code correctness.
+Code coverage refers to how many lines of code percent-wise are currently being tested.
+For example, if I write a test with 70\% test coverage it means that 70\% of the lines of code I wrote are being tested.
+Time simply means how long it takes to write or run the actual tests.
+Code correctness refers to how many errors are found within code after it is considered finished.
+These three attributes are popular to use because they can be quantitatively measured and are considered important within the testing community.
 
-It is important to note that there is no one standard quantitative measurement for code quality because it is a qualitative measurement.
-Because of this many different types of measurement are used to attempt to prove that one testing methodology is better then another.
-Although there are many different attributes that applicable for arguing the superiority of one methodology over another this paper will only focus on four attributes that are commonly found throughout research; code coverage, time efficiency, customer satisfaction, and code correctness.
-\mycomment{I'm not sure this paragraphs approach is the best way to get the information I want across to the reader, I probably should probably put more detail into each of my four attributes}
+\subsection{Waterfall Development and Testing}
 
-\mycomment{may need to add a small bit about functional testing vs unit testing for section 4}
+From as early as the 1970s a popular software development model was the waterfall model where software was developed in a series of phases.
+The waterfall method has been a popular development model because it is considered simple to implement correctly and is also considered to be time efficient because completing previous phases helps latter phases become more efficient.
+The first phase in the standard waterfall model was the requirements phase where requirements were set by the customer or design company.
+Next the design phase occurred where the product was designed which was then followed by the implementation phase where the product was actually created.
+The final phase of the development phase was then the verification phase where testing and debugging occurred \mycomment{wikipedia waterfall model}.
+Due to the fact that these processes often put testing at the end of development, a certain type of testing method, the test-last method, became very popular to use with the waterfall development method.
 
-\subsection{Testing Methods}
-In this paper we will be mainly focusing on two different testing domains, waterfall/test last testing and TDD or test first a popular agile testing method.
+Test last testing is currently a popular testing method and is usually the first testing method that people tend to implement.
+Test Last testing is the practice of writing tests after code has been written that check the functionality of their code.
+These tests are then used to check the written code for any potential errors which are then fixed by the developer until no errors are found by the tests.
+Test last testing is considered simple to implement and can allow the programmer to focus on writing tests based on areas of his code that he may be concerned about.
 
-Waterfall or test last testing is currently a very popular testing method and is usually the first way of testing that people implement.
-This method follows the basic principle of a person first writes their code and then proceeds after the code is considered initially implemented to write tests that check the functionality of their code.
-The developer then runs the test against the code and changes the code until the code passes the developers tests.
-This testing is often used with plan based development or testing on the fly because it allows the developer to only have to write tests about parts of the code he or she is worried about.
-\mycomment{This paragraph needs some serious clean up and much more detail on the topic but for now it gets the basic point across}
+\subsection{Agile Oriented Development and Testing}
 
-TDD or Test first testing is another popular testing method and is primarily used in the agile community.
-This follows the basic principles that one should actually write tests before code and tailor the code to the tests instead of tailoring the tests to the code.
-The idea behind this testing style is that by doing this one can write their codes objectives in tests so that your code becomes tailored to your objectives instead of objectives being tailored to your code.
-\mycomment{again this paragraph needs some serious clean up and detail but for now it gets the basic point across}
+In the late 1990s some software developers started to criticize the sequence-oriented waterfall model, complaining that it was too brittle and inflexible to meet the demands of the standard customer.
+In response to these critiques of the waterfall model a new model for developing software emerged, the agile development model.
+This new development model, based on the tenets of the agile manifesto \mycomment{cite manifesto here!}, promoted the idea that all actions of development should not occur in an ordered sequence but in a series of time-boxed iterations where in each iteration setting requirements, design, implementation and testing would occur simultaneously.
+The goal of each iteration is to produce a demonstrable working product to show a customer and receive feedback on the product.
+Some current development practices that are considered agile are Extreme Programming and Scrum.
 
-\mycomment{things to add to this section}
+Due to the changes in Agile programming test-last testing was pushed aside in favor of a new style known as test-first testing.
+In 1996 \mycomment{might be 2001} with the release of the agile development practice Extreme Programming, the idea of test first testing, implemented in test driven development, started to become popular for the first time ~\cite{Hammond:2012}.
+Test-first testing is the practice of writing tests before code has been written and then writing code to make the tests pass.
+It should be noted that since tests occur before development test-first testing tends to be heavily linked to development methods and thus the most common test-first models also include development elements as well.
 
--Add a brief history to each section
-
--Add justification for why TDD is agile or why TDD is used as an agile practice
-
--Use a more specific TDD example by using extreme programming
-
--write a brief conclusion paragraph containing:
-
--That TDD and waterfall testing are vague terms and not specific
-
--Mention that there will also be a focus Evolved TDD forms briefly(BDD, ATDD)
+The most well known and most used test-first model is that of test driven development or TDD for short.
+In its current use in the field the phrase TDD is mostly used as a blanket term for any sort of development practice that uses test-first testing to drive code development.
+That being said it is important to note that although TDD has been turned into a vague term in this paper we will focus more on its original specific definition.
 

### 2013-11-08 Paper_Draft.tex
Commiting paper.

@@
 
-\title{Agile Testing Four Page Draft}
+\title{The Current State of Test-First vs. Test-Last Testing Draft}
 
@@
 }
-
 \maketitle
 \begin{abstract}
-\mycomment{Needs to be written!}
-
-\mycomment{passages or ideas worth putting into my abstract below}
 
 When it comes to software development perhaps one of the most important and time consuming processes is that of software testing.
-In fact, some early studies on software testing estimated that software testing could consume fifty percent or more of development costs for a product.
-
-Currently there is much debate in the testing world about whether or not test first testing or test last testing is superior and what advantages one type of testing may offer over the other.
-The goal of this paper is to therefore attempt to resolve this debate as much as possible by analyzing research data in the field concerning the advantages and disadvantages of each of these testing methods.
-
-Due to these issues new methods are starting to appear in the agile community that contain some of TDD's main tenets but have shifted enough away from TDD at some fundamental level that they are starting to receive their own names and classifications.
+In fact, early studies on software testing estimated that software testing could consume fifty percent or more of development costs for a product.
+Because of this, the ability to optimize testing to reduce testing costs can be very valuable.
+In this paper  we compare two popular methods, test-last testing, often used in waterfall software development methods, and test-first testing, often used in Agile test driven development software development methods, by reviewing recent studies on the subject.
+In this review we discuss the possible benefits of test-first and test-last testing and possible problems with the current data comparing these two testing methods.
+After that, we explore other methods in test-first testing besides test driven development, such as behavior driven development, in an attempt to find a better test-first testing model.
+In the end we discuss our results and potential future studies to help clarify current data.
 \end{abstract}
@@
 When it comes to software development, perhaps one of the most important and time consuming processes is software testing.
-In fact, some early studies on software testing estimated that it could consume fifty percent or more of development costs for a product ~\cite{Bertolino;2007}.
-because of this businesses over time have become increasingly interested in attempting to optimize testing to reduce development costs.
+In fact, some early studies on software testing estimated that it could consume fifty percent or more of development costs for a product ~\cite{Bertolino:2007}.
+because of this, businesses over time have become increasingly interested in attempting to optimize testing to reduce development costs.
 
-Although there are many testing methods that currently exist two particular testing methods are currently very popular: test-last testing and test-first testing.
-Test-last testing, used mostly in process oriented development, is a testing method where testing is done after the software code has been written to ensure that the code works correctly.
-Test-first testing, used mostly in interval or sprint oriented development, is a testing method where testing is done before the software code is written to ensure that the future code meets certain requirements.
+Although there are many testing methods that currently exist, two particular testing methods are currently very popular: test-last testing and test-first testing.
+Test-last testing, used mostly in process oriented or waterfall development, is a testing method where testing is done after the software code has been written to ensure that the code works correctly.
+Test-first testing, used mostly in interval or agile oriented development, is a testing method where testing is done before the software code is written to ensure that the future code meets certain requirements.
 
@@
 The goal of this paper is to attempt to give an overview of the current state of this debate by analyzing current research data concerning the advantages and disadvantages of each of these testing methods.
-Also because research often lags behind current implementations and the field of test first testing is currently changing due to its relatively new implementation this paper will also explore where new test first methods are going and compare them to the research given.
+Because research often lags behind current implementations and the field of test-first testing is currently changing due to its relatively new implementation, this paper will also explore where new test-first methods are going and compare them to the research given.
 This paper will therefore discuss the advantages and disadvantages of test first testing versus test last testing and explore new emerging test-first methodologies in an attempt to determine the applicability of each methodology.
@@
 In section two of the paper we will discuss what software testing is and current software development models with their supporting testing methods: test-first and test-last testing.
-In section three we will provide an analysis of the data explaining the potential advantages and disadvantages of test first testing compared to test last testing.
-In section four we will go over new test first based testing methodologies and in section five we will state our conclusions and suggestions for further research in the field.
+In section three we will provide an analysis of the data explaining the potential advantages and disadvantages of test-first testing compared to test-last testing.
+In section four we will go over new test-first based testing methodologies and in section five we will state our conclusions and suggestions for further research in the field.
 
@@
 When it comes to judging the effectiveness of certain testing methods it is important to note that there is no one standard quantitative measurement.
-Because of this many different types of measurement are used to attempt to prove that one testing methodology is better then another.
-Although there are many different attributes that are used for arguing the superiority of one method over another we will only focus on three attributes that are commonly found throughout research; code coverage, time, and code correctness.
+Because of this, many different types of measurement are used in an attempt to prove that one testing methodology is better then another.
+Although there are many different attributes that are used for arguing the superiority of one method over another, we will only focus on three attributes that are commonly found throughout research; code coverage, testing time, and code correctness.
 Code coverage refers to how many lines of code percent-wise are currently being tested.
 For example, if I write a test with 70\% test coverage it means that 70\% of the lines of code I wrote are being tested.
-Time simply means how long it takes to write or run the actual tests.
+Testing time simply means how long it takes to write the actual tests.
 Code correctness refers to how many errors are found within code after it is considered finished.
@@
 Test last testing is currently a popular testing method and is usually the first testing method that people tend to implement.
-Test Last testing is the practice of writing tests after code has been written that check the functionality of their code.
+Test Last testing is the practice of writing tests after code has been written to check the functionality of the written code.
 These tests are then used to check the written code for any potential errors which are then fixed by the developer until no errors are found by the tests.
@@
 In response to these critiques of the waterfall model a new model for developing software emerged, the agile development model.
-This new development model, based on the tenets of the agile manifesto \mycomment{cite manifesto here!}, promoted the idea that all actions of development should not occur in an ordered sequence but in a series of time-boxed iterations where in each iteration setting requirements, design, implementation and testing would occur simultaneously.
+This new development model, based on the tenets of the agile manifesto \mycomment{cite manifesto here!}, promoted the idea that all actions of development should not occur in an ordered sequence but in a series of time-boxed iterations where in each iteration setting requirements, product design, product implementation and product testing would occur simultaneously.
 The goal of each iteration is to produce a demonstrable working product to show a customer and receive feedback on the product.
@@
 
-Due to the changes in Agile programming test-last testing was pushed aside in favor of a new style known as test-first testing.
+Due to the changes in Agile programming, test-last testing was pushed aside in favor of a different style known as test-first testing.
 In 1996 \mycomment{might be 2001} with the release of the agile development practice Extreme Programming, the idea of test first testing, implemented in test driven development, started to become popular for the first time ~\cite{Hammond:2012}.
@@
 The most well known and most used test-first model is that of test driven development or TDD for short.
-In its current use in the field the phrase TDD is mostly used as a blanket term for any sort of development practice that uses test-first testing to drive code development.
+In its current use in the field, the phrase TDD is mostly used as a blanket term for any sort of development practice that uses test-first testing to drive code development.
 That being said it is important to note that although TDD has been turned into a vague term in this paper we will focus more on its original specific definition.
+\mycomment{add more once we get to section 3}
 
-\section{Analysis of TDD}
-\subsection{TDD Stances and Data}
-Currently, TDD is the most popular testing method in the Agile community.
-Many supporters of TDD claim that it reduces overall time spent on a product, improves test coverage, and increases overall code quality.
-More importantly there exists data to back up some of these claims.
-for example, in the experiment documented by ~\cite{Lemos:2012}, it was noted that on average TDD methodologies increased code coverage by 40\% compared to conventional testing methods.
-Another study, documented in ~\cite{Hellman:2012} also suggested some positive traits about TDD by claiming some research showed that TDD practices could reduce effort by up to 27\%.
-Studies in articles ~\cite{Hellman:2012, Hammond:2012, Kettunen:2010} acknowledged multiple studies that empirically showed at least a marginal increase in product quality when TDD was implemented.
-By looking at this data alone it would be relatively easy to see why TDD has such a positive image in industry and academia.
+\section{Research Data on Test-First vs Test-Last Testing}
+In this section we attempt to attempt to obtain useful conclusions between test-first and test-last testing by observing current research articles.
+In order to accomplish this we will first start with a data subsection that will go over the three main studies that will be used to draw our conclusions.
+Next we will have an analysis subsection that will summarize the results of the study in terms of our three main empirical measurements: code coverage, code correctness, and total development time.
+We will then have a discussion subsection that will take about potential issues with summarizing the data given in the research articles and then we shall proceed with a conclusion subsection where we will talk about potential conclusions we can draw from the data.
 
-That being said, TDD also has many detractors as well who would tell you that TDD actually increases time spent on a product while not increasing overall code quality.
-This side also has a fair bit of research backing its opinion.
-For example, in ~\cite{Lemos:2012} the same study that showed that TDD increased test coverage by 40\%, it was noted that TDD took significantly more time to implement then conventional testing and did not actually increase code quality.
-Also, the same study that found research showing the TDD could reduce effort by up to 27\% found other research that suggested that TDD actually increased the effort by two fold ~\cite{Hellman:2012}.
-It is also worth noting that articles ~\cite{Hammond:2012, Kettunen:2010} also have data and reports arguing that TDD takes more time then conventional testing.
-What is perhaps troubling about the data given here is that most of it directly contradicts the data given in the first paragraph.
-This suggests an interesting paradox where even though data given on TDD is statistically significant it is decidedly inconclusive.
+\subsection{Data}
+ In 2012 Lemos et.
+al ~\cite{Lemos:2012} conducted a study on sixth semester students to see if test-first testing would increase code coverage, increase testing code size, increase code correctness, and/or decrease in total development time in auxiliary functions (functions with 10-200 lines of code) compared to test-Last testing.
+In this study 39 sixth semester students, knowledgeable in testing techniques and participants of two 100 minute long test first modules, were asked to complete coding challenges over two sessions.
+In the first session, half the students implemented testing-first testing while the other half implemented test-last testing to help solve the problem.
+In the second section the students were asked to switch roles and given a different coding challenge.
+In order to reduce domain knowledge bias amongst the coding challenges three different challenge question types were handed out at each session.
+The results of the study, which were analyzed for statistical significance using the Wilcoxon/Mann-Whitney non-parametric signed-rank paired test, found that test-first testing significantly increased code coverage, testing code size and code development time with no significant increase in code correctness.
 
-\subsection{Analysis of Conflicting Data}
-Even though the data surrounding TDD is inconclusive as a whole it should still be noted that information can still be extracted from it.
-This is especially true because many of the results were statically significant meaning that some sort of meaningful data was obtained.
-Because most of the data was statistically sound it may be worth considering not that the data was faulty but perhaps that the studies contained potentially key differences from one another.
-Thus it may be useful to us to try to understand why the data may be so conflicting as it may lead to useful insights about TDD indirectly.
+In 2010 Kollanus ~\cite{Kollanus:2010} conducted a literary review of TDD.
+Using standard literary review methods, Kollanus found and reviewed forty different experiments in scientific journals, magazines, and conference proceedings that provided empirical evidence for TDD.
+In the review Kollanus found that the most used measurements in these experiments could be clumped into three measurements: external code quality, internal code quality, and productivity.
+In the study Kollanus observed that there was weak support for improved external code quality, very weak support for internal quality, and support for an increased amount of development time.
 
-One of the largest issues for summarizing the effectiveness of TDD is the field itself is very broad.
-It is important to note that while I specifically defined the original ideas and methodologies of the original idea of TDD, TDD itself has become an umbrella term that has been used to describe a massive array of different testing practices.
-This diversity may help explain why we are getting differing results because the TDD practice in one study may be extraordinary different from the TDD practice in another study.
-It is worth noting that in ~\cite{Hammond:2012} a major issue with many TDD research papers was the absence of how the TDD process was implemented.
-This is a major issue in studying the field of TDD because it is hard to tell if people used similar or different methods of TDD to obtain their results.
+In 2012 Hellmann et.
+al conducted a summary on the current status of research of TDD.
+Within Hellmann’s background there is a mention of a summary study produced by Jeffries and Mitnik that found in general that TDD largely resulted in an increase in quality, but one study they identified showed instead that TDD resulted in a strong negative impact on quality.
+Additionality while they showed that TDD could reduce the amount of effort required by up to 27\%, most studies found an increase in effort of up to 100\%.
 
-Another issue for understanding the effectiveness of TDD that many TDD researchers have run into is that TDD is actually a difficult process to implement correctly.
-There were many testimonies throughout various articles ~\cite{Hammond:2012, Hellman:2012, Kettunen:2010} of industry professionals finding the TDD process to be much more difficult to implement correctly compared to traditional testing frameworks.
-Also, many articles on the analysis of TDD at some point brings up the topic of the complexity of TDD.
-This acknowledgement of complexity is important as it may suggest a result bias in TDD experiments based on the skill level of the participants as less experienced participants may be significantly less likely to implement TDD in an efficient or successful manner while more experienced participants may have more success.
-In one study at a college, although not significantly analyzed due to a lack of participants, it was noted that Alumni who used TDD produced overall better quality code compared to their conventional testing Alumni counterparts while current students in the college showed almost no difference in code quality between TDD and conventional test methods ~\cite{Lemos:2012}.
-This potential bias would wreck havoc on TDD data conclusiveness as a whole as the two major participant types for TDD experiments are industry professionals and college students who have vast gaps in experience difference.
+\subsection{Analysis}
+Two of the three studies covered the topic of code coverage.
+In both studies by Kollanus and Lemos et al.
+it was acknowledged that increased test coverage had occurred in test-first testing methods compared to their test-last counterparts.
+All three papers brought up up total development time and noted that test-first testing tended to increase development time, although both Kollanus’s study and Hellmann’s study provide notable contradictory studies to this conclusion.
+All three studies in one way or another also brought up code correctness in the form of external code quality in Kollanus or in the form of code quality in the case of Hellmann et al.
+In this case the results were across the board where Lemos et al.
+stated that there was no code correctness difference between test first and test last whereas Kollanus said there was weak evidence that test-first testing produces more correct code then test-last testing.
+Hellmann et al provided contradictory evidence backing that in most studies test-first testing has shown better code correctness then test last testing but there was a noted study where test-first testing was less code correct then test-last testing.
+In summary, with the studies given, test-first testing produces more code coverage and is likely to take more time to code then test-last testing but due to contradictions in the research given we can not draw a conclusion on code correctness.
 
-Overall it is very hard to draw conclusions on the effectiveness of TDD.
-Out of all the data out there there is really only one data point that seems to be incontrovertible which is the fact that TDD always seems to increase overall code coverage.
-Perhaps another conclusion that can be made is that TDD seems to produce no worse code quality then traditional waterfall testing.
-Also there are some implications from potential confounding factors that not all TDD methods are created equal and experience may play a large role in TDD effectiveness.
-Everything else on the effectiveness of TDD seems to be inconclusive.
+\subsection{Discussion}
+Many summaries of TDD research ~\cite{Hammond:2012, Hellman:2012, Kettunen:2010} , including the literary review done by Kollanus ~\cite{Kollanus:2010}, have noted that one confounding factor to their ability to summarize the research in the given field of test-first testing has been that many studies are contradictory to one another.
+For example in ~\cite{Hellman:2012},  there is a mention of previous summary studies that have had previous issues with conflicting data.
+This is then followed by a series of examples of data that has conflicted in other studies in the past.
+In Kollanus’s literature review ~\cite{Kollanus:2010} Kollanus points out that many articles in her study contradicted other articles which made it hard to make sound conclusions from the research given and may be a confounding factor in her study.
+In most current studies this seems to occur because of two main factors: participant experience and improper study implementation and documentation.
+
+In the case of participant experience there seems to be a skill gap between test first and test last testing.
+In an opinion study by ~\cite{Janzen:2007} it was found that a majority college students said they were not comfortable with and had trouble with comprehending TDD, even after doing a coding exercise with it.
+The majority of students also said that they were much more comfortable implementing test last methods and would prefer to use test last methods in the future.
+Coding professionals, on the other hand, actually preferred test first programming and were much more comfortable with it than students.
+They also stated they felt it was relatively easy to understand while students did not.
+In another opinion study ~\cite{Kettunen:2010}, by Kettunen et al, a common theme throughout the paper was examples of individual testimonies and research conclusions that pointed to test first testing being harder to implement than test last testing.
+Considering that most conducted experiments focus on college students and industry professionals, who have vast experience differences, the fact that test first testing may be harder to implement for college students compared to industry professionals may create differing results between studies.
+This idea is considered in ~\cite{Hammond:2012} and ~\cite{Kettunen:2010}, but no summary papers where found considering this result and comparing student participant experiment results to professional participant experiment results.
+
+The other problem that has plagued summary papers of test first testing is that most studies either lack documentation in key locations or do not effectively handle the participant conformance issue which comes up in TDD studies.
+In the literary review by Kollanus, she acknowledges a frustrating lack of information in many studies on how TDD was implemented.
+Most studies, claimed Kollanus, had maybe one or two lines describing their TDD methodology or just simply claimed that they used TDD.
+Considering that TDD can be implemented in many ways, as discussed earlier in the background, Kollanus states that the lack of documentation for TDD implementation means that, as a reviewer, she has no idea if the test methods used to produce the results were similar or different thus potentially confounding some of her conclusions.
+Another issue similar to this is the conformance issue, brought up in (cite here).
+The conformance issue in TDD testing refers to the fact that few researchers make sure that TDD was implemented correctly or at all by its participants.
+For example, in the study by Lemos et.
+al the researchers acknowledged that one of their confounding factors was that they only asked students to write tests before code, thus a variety of different test first methods might have been used within the study making its data less conclusive.
+Overall the lack of documentation of how researchers specifically planned to implement TDD in their study and their lack of some sort of monitoring of whether the implementation was actually occurring greatly reduces the credibility of the research being done and allows for confounding factors to potentially occur.
+
+\subsection{Conclusion}
+
+Due to current issues in contradictory data it is hard to make solid conclusions about the advantages and disadvantages of using test first testing instead of test last testing.
+What can be done instead is we can consider certain trends in research that can suggest potential answers.
+Most studies and summaries seem to agree that test-first testing tends to produce more test coverage compared to test-last testing and that test first testing tends to take longer to write tests compared to test first testing.
+Two of the studies we looked at, the experiment by lemos and the literary review by Kollanus, point out these two trends.
+Also one or both trends have been noted through multiple papers ~\cite{Kettunen:2010, Hammond:2012, Hellman:2012} with no major contradictions found.
+Another convincing trend that has occurred is that in general test first testing seems to be harder to implement than test last testing as mentioned earlier.
+Although these trends seem to be clear, the rest seem to be more muddled.
+Code correctness for example seems to have no clear trend as multiple studies have reached contradictory conclusions to other studies.
+Another hot button topic that fails to show a trend one way or the other is the amount of time spent simply developing code as this again seems to have results across the board.
+From these trends we can consider that test first testing overall probably takes more time to write, increases code coverage, and is harder to implement than test-last testing and that other advantages or detriments may exist but they are unclear in current research.
+
 \section{evolutions to TDD}
-\mycomment{I feel that this is a much needed and very interesting section for exploring Agile testing but I am becoming increasingly concerned about how well this section may be able to tie into the rest of the paper.
-I would like to use this section as a useful resource for seeing where TDD is going/ what problems TDD developers are trying to address/ see if any of these new spin offs may solve problems that are coming up in research listed in section 3.
-Any input on how to tie this section into the paper or how to better reach the goals given above would be useful.}
+\mycomment{this section is currently a work in progress}
 
@@
 Due to these issues new methods are starting to appear in the agile community that contain some of TDD's main tenets but have shifted enough away from TDD at some fundamental level that they are starting to receive their own names and classifications.
-Although there are many of these spin-offs in TDD for this paper we chose to focus on three more popular spin offs, acceptance test driven development, Behavior driven development, and Agile Specification Driven Development.
+Although there are many of TDD spin-offs, for this paper we chose to focus on three more popular spin offs, acceptance test driven development, Behavior driven development, and Agile Specification Driven Development.
 
-\subsection{acceptance test driven development}
+\subsection{Test Driven Development}
+Test driven development, as mentioned earlier is currently the most widely used test first testing and design methodology.
+Although in the previous cases we have been using TDD as a vague term in this section we will focus on the the original TDD methodology introduced in Extreme Programming.
+In the original TDD methodology the developer repeats a step of processes to develop his code.
+The process begins once all current tests pass or the new code has just been started.
+When this occurs the programmer writes a new failing test that tests the simplest functionality the programmer wishes to add to their code.
+Once the test has been written and the test fails, the programmer then writes the minimal amount of code to make the test pass.
+After this step the programmer proceeds to refactor any clutter within his solution and then repeats the process until the code is complete.
+
+Although TDD is a very popular style of test first testing, it should be noted that a perceived weakness to the TDD methodology exists.
+The main issue many people have with test driven development is that it fails to suggest a style of testing and how to best define an independent functionality in ones code plan.
+In the study ~\cite{Kettunen:2010} (cite others here as needed) people noted that TDD was difficult because they had a hard time deciding how to implement their tests effectively.
+Another related issue noted in ~\cite{Kettunen:2010} was that some individuals found it hard to figure out what needed to specifically be tested thus some found they wrote an overabundance of tests causing the process to become inefficient and thus take more time.
+Others found that focusing on the simplest functions first left serious potential problems to the end of development causing a serious time crunch problem for product release.
+Due to these issues with defining functionality and other issues within TDD new methods for implementing test first testing are being designed and implemented in an attempt to solve these problems.
+\subsection{Acceptance Test Driven Development}
 
@@
 \mycomment{need to do research here, specifically it would be useful to do some research on DbC}
+
+\subsection{section conclusion}
+\mycomment{the goal here is to have a discussion over which TDD spin off might be most promising and for what reasons}
+
 \section{conclusion}
-In conclusion with the current information available to us there is no obviously superior testing methods between TDD or test first ideologies and waterfall or test last ideologies.
-That being said it is worth noting that their are still differences between the two testing methods and that although there is not a clear overall advantage between the two languages certain traits make make one method of testing better then another method of testing in a specific scenario.
-For example when dealing with new and relatively inexperienced programmers the test last methodology may be more efficient then the test first methodology.
-Another example may be that you believe code coverage is an important necessity for your product, in this case you would probably want to use test first over test last to test your code.
-Therefore until research on the subject becomes more clear it would be wise to consider both testing practices as effective and potentially useful for any project you may choose to undertake.
+\mycomment{this section is currently a work in progress}
+\subsection{Conclusion intro}
 
-\mycomment{this paragraph is a rough idea of what might be useful conclusions to draw from section three}
+In this paper we talked about two common test methods: test-first testing, often found in waterfall development models and test-last testing, often found in agile development models.
+In section two we defined these test methods as well as provided background information to the reader.
+In section three we explored recent research into the comparison of the two different testing methods and explored potential research problems with the current data.
+In particular we found that data was contradictory but we still were able to tentatively conclude that test-first testing increases test coverage, increases time taken to test, and is harder to implement than test-last testing.
+In this section we also suggested two potential reasons for the contradictory data: different participant experience levels and poor documentation/implementation of specific test first methods.
+In section four we explored different implementations of test first testing.
 
+\subsection{Discussion of Implications of Section 3}
+\mycomment{stubbed for now}
+
+\subsection{Discussion of Implications of Section 4}
+\mycomment{stubbed for now}
+\subsection{Suggestion for Further Research}
 For people who are looking to advance the field of TDD it would be potentially very interesting to see how the new TDD spin-offs such as Behavior Driven Development and Agile Specification-Driven Development compare to that of a well documented TDD process like the one suggested for Extreme Programming.
-Another thing that could also help advance the field is to do research comparing the effectiveness of waterfall testing and TDD testing between groups with notably different experience levels as some research tentatively points that this may be a relevant variable.
+Another useful study that could be implemented would be to compare the results of studies thing that could also help advance the field is to do research comparing the effectiveness of waterfall testing and TDD testing between groups with notably different experience levels as some research tentatively points that this may be a relevant variable.
 

### 2013-11-09 Paper_Draft.tex
Draft for friday night.

@@
 
-When it comes to software development perhaps one of the most important and time consuming processes is that of software testing.
-In fact, early studies on software testing estimated that software testing could consume fifty percent or more of development costs for a product.
+When it comes to software development, perhaps one of the most important and time consuming processes is that of software testing.
+In fact, early studies on software testing estimated that it could consume fifty percent or more of development costs for a product.
 Because of this, the ability to optimize testing to reduce testing costs can be very valuable.
-In this paper  we compare two popular methods, test-last testing, often used in waterfall software development methods, and test-first testing, often used in Agile test driven development software development methods, by reviewing recent studies on the subject.
+In this paper  we compare two popular methods, test-last testing, often used in waterfall software development methods, and test-first testing, often used in Agile test driven software development methods, by reviewing recent studies on the subject.
 In this review we discuss the possible benefits of test-first and test-last testing and possible problems with the current data comparing these two testing methods.
@@
 In fact, some early studies on software testing estimated that it could consume fifty percent or more of development costs for a product ~\cite{Bertolino:2007}.
-because of this, businesses over time have become increasingly interested in attempting to optimize testing to reduce development costs.
+Because of this, software developers have become increasingly interested in attempting to optimize testing to reduce development costs.
 
@@
 
-Currently, there is much debate in the testing community about whether or not test first or test last testing is superior.
+Currently, there is much debate in the testing community about whether or not test-first or test-last testing is superior.
 The goal of this paper is to attempt to give an overview of the current state of this debate by analyzing current research data concerning the advantages and disadvantages of each of these testing methods.
@@
 
-The rest of the paper is divided into four sections which will discuss the specific ideas above in more detail.
-In section two of the paper we will discuss what software testing is and current software development models with their supporting testing methods: test-first and test-last testing.
+The rest of the paper is divided into four sections.
+In section two of the paper we discuss what software testing is and current software development models with their supporting testing methods: test-first and test-last testing.
 In section three we will provide an analysis of the data explaining the potential advantages and disadvantages of test-first testing compared to test-last testing.
@@
 Because of this, many different types of measurement are used in an attempt to prove that one testing methodology is better then another.
-Although there are many different attributes that are used for arguing the superiority of one method over another, we will only focus on three attributes that are commonly found throughout research; code coverage, testing time, and code correctness.
+Although there are many different attributes that are used for arguing the superiority of one testing method over another, we will focus on three attributes that are commonly found throughout research; code coverage, total development time, and code correctness.
 Code coverage refers to how many lines of code percent-wise are currently being tested.
-For example, if I write a test with 70\% test coverage it means that 70\% of the lines of code I wrote are being tested.
-Testing time simply means how long it takes to write the actual tests.
+For example, if I write a test with 70\% test coverage it means that 70\% of the lines of code that were wrote are being tested.
+Total development time refers to how long it takes to finish a development project.
 Code correctness refers to how many errors are found within code after it is considered finished.
@@
 
-From as early as the 1970s a popular software development model was the waterfall model where software was developed in a series of phases.
-The waterfall method has been a popular development model because it is considered simple to implement correctly and is also considered to be time efficient because completing previous phases helps latter phases become more efficient.
-The first phase in the standard waterfall model was the requirements phase where requirements were set by the customer or design company.
-Next the design phase occurred where the product was designed which was then followed by the implementation phase where the product was actually created.
-The final phase of the development phase was then the verification phase where testing and debugging occurred \mycomment{wikipedia waterfall model}.
+From as early as the 1970s, a popular software development model was the waterfall model, where software was developed in a series of phases.
+The waterfall development model is used because it is simple to implement correctly and is also considered time efficient because completing previous phases helps later phases become more efficient.
+The first phase in the standard waterfall model is the requirements phase where requirements are set by the customer or design company.
+Next the design phase occurs in which the product is designed and then the product is built in the implementation phase.
+The final phase is the verification phase where testing and debugging occurred \mycomment{wikipedia waterfall model}.
 Due to the fact that these processes often put testing at the end of development, a certain type of testing method, the test-last method, became very popular to use with the waterfall development method.
 
-Test last testing is currently a popular testing method and is usually the first testing method that people tend to implement.
-Test Last testing is the practice of writing tests after code has been written to check the functionality of the written code.
-These tests are then used to check the written code for any potential errors which are then fixed by the developer until no errors are found by the tests.
-Test last testing is considered simple to implement and can allow the programmer to focus on writing tests based on areas of his code that he may be concerned about.
+Test-last testing is currently a popular testing method and is usually the first testing method that people tend to implement.
+Test-last testing is the practice of writing tests after code has been written to check the functionality of the written code.
+These tests are then used by the developer to fix their code until no errors are found by the tests.
+/mycomment{this last sentence may no longer be needed} Test last testing allows the programmer to focus on writing tests based on areas of his code that he may be concerned about.
 
@@
 
-In the late 1990s some software developers started to criticize the sequence-oriented waterfall model, complaining that it was too brittle and inflexible to meet the demands of the standard customer.
-In response to these critiques of the waterfall model a new model for developing software emerged, the agile development model.
-This new development model, based on the tenets of the agile manifesto \mycomment{cite manifesto here!}, promoted the idea that all actions of development should not occur in an ordered sequence but in a series of time-boxed iterations where in each iteration setting requirements, product design, product implementation and product testing would occur simultaneously.
-The goal of each iteration is to produce a demonstrable working product to show a customer and receive feedback on the product.
+In the late 1990s, some software developers started to criticize the sequence-oriented waterfall model, complaining that it was too brittle and inflexible to meet the demands of the standard customer.
+In response to these critiques of the waterfall model, a new model for developing software emerged, the agile development model.
+This new development model, based on the tenets of the agile manifesto \mycomment{cite manifesto here!}, promoted the idea that all actions of development should not occur within an ordered sequence of phases, but in a series of time-boxed iterations where in each iteration developers set requirements, design a product, make it, and test it.
+The goal of each iteration is to produce a demonstrable working product to show a customer and to receive feedback on the product.
 Some current development practices that are considered agile are Extreme Programming and Scrum.
@@
 That being said it is important to note that although TDD has been turned into a vague term in this paper we will focus more on its original specific definition.
-\mycomment{add more once we get to section 3}
+\mycomment{This paragraph will need to be reworked once the TDD section is reworked, it is currently out of date.}
 
 \section{Research Data on Test-First vs Test-Last Testing}
-In this section we attempt to attempt to obtain useful conclusions between test-first and test-last testing by observing current research articles.
-In order to accomplish this we will first start with a data subsection that will go over the three main studies that will be used to draw our conclusions.
-Next we will have an analysis subsection that will summarize the results of the study in terms of our three main empirical measurements: code coverage, code correctness, and total development time.
-We will then have a discussion subsection that will take about potential issues with summarizing the data given in the research articles and then we shall proceed with a conclusion subsection where we will talk about potential conclusions we can draw from the data.
+In this section we will attempt to obtain useful comparisons between test-first and test-last testing by observing current research articles.
+In order to accomplish this, this section starts with a data subsection that goes over three main studies that will be used to draw our conclusions.
+Next, an analysis subsection summarizes the results of the studies in terms of our three main empirical measurements: code coverage, code correctness, and total development time.
+Afterwards a discussion subsection talks about the potential issues with summarizing the data given in the research articles and then the conclusion subsection talks about potential useful comparisons we can draw from the data.
 
 \subsection{Data}
- In 2012 Lemos et.
-al ~\cite{Lemos:2012} conducted a study on sixth semester students to see if test-first testing would increase code coverage, increase testing code size, increase code correctness, and/or decrease in total development time in auxiliary functions (functions with 10-200 lines of code) compared to test-Last testing.
-In this study 39 sixth semester students, knowledgeable in testing techniques and participants of two 100 minute long test first modules, were asked to complete coding challenges over two sessions.
+\mycomment{Reworked the paragraphs on the Kollanus study and the Lemos study in an attempt to better represent data in my paper, also planning on writting paragraphs for two other studies.
+Considering making this its own section due to its increase in size.}
+
+This subsection contains summaries of important studies that will be used later in the paper to draw out important conclusions and comparisons between test-first and test-last testing.
+
+The first research summarized research article  is a 2010 article by Kollanus ~\cite{Kollanus:2010}.
+In this article, Kollanus, using standard literary review methods, found and reviewed forty different experiments in scientific journals, magazines, and conference proceedings that provided empirical evidence about the effectiveness of using TDD compared to test-last development methods.
+In the review Kollanus focuses on three different measurements: external code quality, productivity, and internal code quality.
+In the article, Kollanus describes external code quality as the measurement of the number of passed researcher acceptance tests in a test set when run against code written by participants.
+Since this definition is equivalent to our definition for code-correctness this data can be used for code-correctness comparisons.
+In the study, Kollanus concluded that there was weak support for improved external code quality in TDD methods compared to test-last development methods.
+This conclusion was based on consistent evidence found in case studies that reported better external code quality in TDD methods.
+However, the support was considered weak because there were contradictory results found in some controlled experiments that recorded external code quality.
+Productivity, as defined by Kollanus, is a measurement that measures how much code one writes in comparison to time spent.
+Since this definition is similar to that of total development time's, this data can tentatively be used for total development time comparisons.
+The research viewed by Kollanus heavily suggests that TDD methods are less productive and therefore take more time overall to write.
+This conclusion was drawn from the fact that the majority of studies and experiments showed a decrease in productivity.
+That being said multiple studies seem to contradict the majority.
+Internal code quality describes a wide set of measurements that measure code quality from a development standpoint.
+Although code coverage is one of these measurements and there is a mention that one article found increased code coverage in TDD methods, this sections conclusions will not be discussed as it does not fit with our paper's themes.
+
+In 2012, Hellmann et.
+al conducted a summary on the current status of research of TDD.
+Within Hellmann's background there is a mention of a summary study produced by Jeffries and Mitnik that found in general that TDD largely resulted in an increase in quality, but one study they identified showed instead that TDD resulted in a strong negative impact on quality.
+Additionally while they showed that TDD could reduce the amount of effort required by up to 27\%, most studies found an increase in effort of up to 100\%.
+
+In 2012, Lemos et.
+al ~\cite{Lemos:2012} conducted a study on third year computer science undergraduate students to see if test-first testing would significantly impact code coverage,  code correctness, and/or total development time in auxiliary functions (functions with 10-200 lines of code) compared to test-Last testing.
+In this study 39 third year undergrad computer science students, knowledgeable in testing techniques and participants of two 100 minute long test first modules, were asked to complete coding challenges over two sessions.
 In the first session, half the students implemented testing-first testing while the other half implemented test-last testing to help solve the problem.
@@
 In order to reduce domain knowledge bias amongst the coding challenges three different challenge question types were handed out at each session.
-The results of the study, which were analyzed for statistical significance using the Wilcoxon/Mann-Whitney non-parametric signed-rank paired test, found that test-first testing significantly increased code coverage, testing code size and code development time with no significant increase in code correctness.
+Code coverage was measured by comparing the percentage of tested lines in participants code.
+Total development time was measured by how long it took participants to finish their code.
+Code correctness was measured by running acceptance tests against a participants development code and giving the code a score of 0(all test cases fail), .5(some test cases fail) or 1(all test cases pass).
+Note this style of measuring code correctness has been used in previous studies and is call the \textit{Functional Test Set Success Level} scale.
+The results of the study were analyzed for statistical significance using the Wilcoxon/Mann-Whitney non-parametric signed-rank paired test.
+For the results of code coverage, it was found that code coverage on average was 40\% higher when test-first testing was used.
+On total development time, it was found that test-first code took 12\% longer to write then test-last code.
+Both of these results were statistically significant.
+In terms of code correctness, the only difference found was that the test-last code had one more correct implimentation of code then the test-first code.
+This result was not found to be statistically significant.
 
-In 2010 Kollanus ~\cite{Kollanus:2010} conducted a literary review of TDD.
-Using standard literary review methods, Kollanus found and reviewed forty different experiments in scientific journals, magazines, and conference proceedings that provided empirical evidence for TDD.
-In the review Kollanus found that the most used measurements in these experiments could be clumped into three measurements: external code quality, internal code quality, and productivity.
-In the study Kollanus observed that there was weak support for improved external code quality, very weak support for internal quality, and support for an increased amount of development time.
-
-In 2012 Hellmann et.
-al conducted a summary on the current status of research of TDD.
-Within Hellmann’s background there is a mention of a summary study produced by Jeffries and Mitnik that found in general that TDD largely resulted in an increase in quality, but one study they identified showed instead that TDD resulted in a strong negative impact on quality.
-Additionality while they showed that TDD could reduce the amount of effort required by up to 27\%, most studies found an increase in effort of up to 100\%.
+\mycomment{need to add two paragraphs for the studies by Janzen and Kettenun}
 
@@
 
+\section{Test Driven Development}
+
+\mycomment{currently this section is more of a stub then anything.
+This section should be fleshed out over the weekend.
+Although I am not sure on formatting or precise content yet, many articles have noted that TDD takes longer and is harder to implement because it is unhelpful in determining what parts of the code/problem need to be tested and what is the best way of implementing those tests.}
+Test driven development, as mentioned earlier is currently the most widely used test first testing and design methodology.
+Although in the previous cases we have been using TDD as a vague term in this section we will focus on the the original TDD methodology introduced in Extreme Programming.
+In the original TDD methodology the developer repeats a step of processes to develop his code.
+The process begins once all current tests pass or the new code has just been started.
+When this occurs the programmer writes a new failing test that tests the simplest functionality the programmer wishes to add to their code.
+Once the test has been written and the test fails, the programmer then writes the minimal amount of code to make the test pass.
+After this step the programmer streamlines his solution with the other parts of his code.
+This series of steps repeats until the code is complete.
+
+~\mycomment{this next paragraph is an extremely rough draft that was written before this section became a section and currently makes little to no sense, comments are not needed for this paragraph}
+
+Although TDD is a very popular style of test first testing, it should be noted that a perceived weakness to the TDD methodology exists.
+The main issue many people have with test driven development is that it fails to suggest a style of testing and how to best define an independent functionality in ones code plan.
+In the study ~\cite{Kettunen:2010} (cite others here as needed) people noted that TDD was difficult because they had a hard time deciding how to implement their tests effectively.
+Another related issue noted in ~\cite{Kettunen:2010} was that some individuals found it hard to figure out what needed to specifically be tested thus some found they wrote an overabundance of tests causing the process to become inefficient and thus take more time.
+Others found that focusing on the simplest functions first left serious potential problems to the end of development causing a serious time crunch problem for product release.
+Due to these issues with defining functionality and other issues within TDD new methods for implementing test first testing are being designed and implemented in an attempt to solve these problems.
+
 \section{evolutions to TDD}
@@
 
-\subsection{Test Driven Development}
-Test driven development, as mentioned earlier is currently the most widely used test first testing and design methodology.
-Although in the previous cases we have been using TDD as a vague term in this section we will focus on the the original TDD methodology introduced in Extreme Programming.
-In the original TDD methodology the developer repeats a step of processes to develop his code.
-The process begins once all current tests pass or the new code has just been started.
-When this occurs the programmer writes a new failing test that tests the simplest functionality the programmer wishes to add to their code.
-Once the test has been written and the test fails, the programmer then writes the minimal amount of code to make the test pass.
-After this step the programmer proceeds to refactor any clutter within his solution and then repeats the process until the code is complete.
-
-Although TDD is a very popular style of test first testing, it should be noted that a perceived weakness to the TDD methodology exists.
-The main issue many people have with test driven development is that it fails to suggest a style of testing and how to best define an independent functionality in ones code plan.
-In the study ~\cite{Kettunen:2010} (cite others here as needed) people noted that TDD was difficult because they had a hard time deciding how to implement their tests effectively.
-Another related issue noted in ~\cite{Kettunen:2010} was that some individuals found it hard to figure out what needed to specifically be tested thus some found they wrote an overabundance of tests causing the process to become inefficient and thus take more time.
-Others found that focusing on the simplest functions first left serious potential problems to the end of development causing a serious time crunch problem for product release.
-Due to these issues with defining functionality and other issues within TDD new methods for implementing test first testing are being designed and implemented in an attempt to solve these problems.
 \subsection{Acceptance Test Driven Development}

### 2013-11-23 Paper_Draft.tex
pushing stuff for funsies

--- 2013-11-08 Paper_Draft.tex
@@
 
-\title{The Current State of Test-First vs. Test-Last Testing Draft}
+\title{An Overview of the Current State of Test-First vs. Test-Last Debate}
 
@@
 
-When it comes to software development perhaps one of the most important and time consuming processes is that of software testing.
-In fact, early studies on software testing estimated that software testing could consume fifty percent or more of development costs for a product.
+When it comes to software development, perhaps one of the most important and time consuming processes is that of software testing.
+In fact, early studies on software testing estimated that it could consume fifty percent or more of development costs for a product.
 Because of this, the ability to optimize testing to reduce testing costs can be very valuable.
-In this paper  we compare two popular methods, test-last testing, often used in waterfall software development methods, and test-first testing, often used in Agile test driven development software development methods, by reviewing recent studies on the subject.
+In this paper  we compare two popular methods, test-last testing, often used in waterfall software development processes, and test-first testing, often used in Agile test driven software development methods, by reviewing recent studies on the subject.
 In this review we discuss the possible benefits of test-first and test-last testing and possible problems with the current data comparing these two testing methods.
@@
 % A category with the (minimum) three required fields
-\category{H.4}{Information Systems Applications}{Miscellaneous}
-%A category including the fourth, optional field follows...
-\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]
-
-\terms{Delphi theory}
 
@@
 When it comes to software development, perhaps one of the most important and time consuming processes is software testing.
-In fact, some early studies on software testing estimated that it could consume fifty percent or more of development costs for a product ~\cite{Bertolino:2007}.
-because of this, businesses over time have become increasingly interested in attempting to optimize testing to reduce development costs.
+In fact, some early studies on software testing estimated that it could consume fifty percent or more of the development costs for a product ~\cite{Bertolino:2007}.
+Because of this, software developers have become increasingly interested in attempting to optimize testing to reduce development costs.
 
-Although there are many testing methods that currently exist, two particular testing methods are currently very popular: test-last testing and test-first testing.
-Test-last testing, used mostly in process oriented or waterfall development, is a testing method where testing is done after the software code has been written to ensure that the code works correctly.
-Test-first testing, used mostly in interval or agile oriented development, is a testing method where testing is done before the software code is written to ensure that the future code meets certain requirements.
+Although there are many testing methods that  exist, two particular testing methods are currently very popular: test-last testing and test-first testing.
+Test-last testing, used mostly in process oriented or waterfall development, is a testing method where testing is done after software is written to ensure that the software is working as intended.
+Test-first testing, used mostly in interval or agile oriented development, is a testing method where tests are written before the software being tested is written to ensure that the code to be written meets certain requirements.
 
-Currently, there is much debate in the testing community about whether or not test first or test last testing is superior.
-The goal of this paper is to attempt to give an overview of the current state of this debate by analyzing current research data concerning the advantages and disadvantages of each of these testing methods.
-Because research often lags behind current implementations and the field of test-first testing is currently changing due to its relatively new implementation, this paper will also explore where new test-first methods are going and compare them to the research given.
-This paper will therefore discuss the advantages and disadvantages of test first testing versus test last testing and explore new emerging test-first methodologies in an attempt to determine the applicability of each methodology.
+Recently, there has been much debate in the testing community about whether or not test-first or test-last testing is superior.
+The goal of this paper is to attempt to give an overview of the current state of this debate by analyzing current research data concerning the advantages and disadvantages of each testing method.
+Because research often lags behind current implementations, and the field of test-first testing is currently changing due to its relatively new implementation, this paper will also explore new test-first methods.
+This paper will discuss the advantages and disadvantages of test-first testing versus test-last testing and explore new test-first methodologies in an attempt to determine the applicability of each methodology.
 
-The rest of the paper is divided into four sections which will discuss the specific ideas above in more detail.
-In section two of the paper we will discuss what software testing is and current software development models with their supporting testing methods: test-first and test-last testing.
-In section three we will provide an analysis of the data explaining the potential advantages and disadvantages of test-first testing compared to test-last testing.
-In section four we will go over new test-first based testing methodologies and in section five we will state our conclusions and suggestions for further research in the field.
+The paper is divided into five sections.
+In Section II we discuss what software testing is and current software development models with their supporting testing methods: test-first and test-last testing.
+In Section III we will provide an analysis of the data explaining the potential advantages and disadvantages of test-first testing compared to test-last testing.
+In Section IV we discuss issues of using test driven development to implement test-first testing.
+In Section V we will go over new testing methodologies based on test-first development.
+In Section VI we will provide conclusions and suggestions for further research in the field.
 
@@
 \subsection{Software Testing}
-Software testing, simply defined, is a branch of software engineering that entails a series of practices meant to either identify potential malfunctions or demonstrate functionality in a software system ~\cite{Bertolino:2007}.
-Software testing can be as simple as running a program to see its results or can be as complex as writing code to simulate scenarios in the real world.
+Software testing, simply defined, is a branch of software engineering that uses a series of practices meant to either identify potential malfunctions or demonstrate functionality in a software system ~\cite{Bertolino:2007}.
+Software testing can be as simple as running a program to see if its results look correct or can be as complex as writing code to simulate scenarios in the real world.
 
-When it comes to judging the effectiveness of certain testing methods it is important to note that there is no one standard quantitative measurement.
-Because of this, many different types of measurement are used in an attempt to prove that one testing methodology is better then another.
-Although there are many different attributes that are used for arguing the superiority of one method over another, we will only focus on three attributes that are commonly found throughout research; code coverage, testing time, and code correctness.
-Code coverage refers to how many lines of code percent-wise are currently being tested.
-For example, if I write a test with 70\% test coverage it means that 70\% of the lines of code I wrote are being tested.
-Testing time simply means how long it takes to write the actual tests.
+When comparing different testing methods there is no one standard quantitative measurement that determines which method is superior.
+Because of this, many different types of measurement are used to argue that one testing method is better then another.
+In this paper we will focus on three attributes that are commonly found throughout research: code coverage, total development time, and code correctness.
+These three attributes are popular because they can be quantitatively measured and are considered important within the testing community.
+
+Code coverage refers to the percentage of lines of production code that are executed when a set of tests are run.
+For example, if I write a set of tests that has 70\% test coverage, it means that 70\% of the lines of production code were executed when the tests were run.
+Total development time refers to how long it takes to finish a development project.
 Code correctness refers to how many errors are found within code after it is considered finished.
-These three attributes are popular to use because they can be quantitatively measured and are considered important within the testing community.
+This is often measured by running a very large all encompassing test set against the participants code in the attempt to find cases where their code fails to produce the desired result.
 
 \subsection{Waterfall Development and Testing}
+\mycomment{present tense this paragraph}
+A popular software development model that was developed in the 1970's is the waterfall software development model, where software is developed in a series of phases.
+The waterfall development model is popular because it is simple to implement correctly and is time efficient.
+The first phase in the standard waterfall model is the requirements phase, where requirements are set by the customer or design company.
+Next the design phase occurs in which the product is designed.
+The product is then built in the implementation phase.
+The final phase is the verification phase where testing and debugging occur ~\cite{wiki:xxx}.
+The phases were set up in this manner to have previous phases make later phases simpler to complete.
+Due to the fact that these processes often put testing at the end of development, a certain type of testing method, test-last testing, was the only testing method that made sense to use.
+http://start.fedoraproject.org/
+Test-last testing is currently a popular testing method and is usually the first testing method that people tend to implement.
+Test-last testing is the practice of writing tests after code has been written to check the functionality of the written code.
+These tests are then used by the developer to fix their code until no further errors are found by the tests.
 
-From as early as the 1970s a popular software development model was the waterfall model where software was developed in a series of phases.
-The waterfall method has been a popular development model because it is considered simple to implement correctly and is also considered to be time efficient because completing previous phases helps latter phases become more efficient.
-The first phase in the standard waterfall model was the requirements phase where requirements were set by the customer or design company.
-Next the design phase occurred where the product was designed which was then followed by the implementation phase where the product was actually created.
-The final phase of the development phase was then the verification phase where testing and debugging occurred \mycomment{wikipedia waterfall model}.
-Due to the fact that these processes often put testing at the end of development, a certain type of testing method, the test-last method, became very popular to use with the waterfall development method.
+\subsection{Agile Development and Testing}
 
-Test last testing is currently a popular testing method and is usually the first testing method that people tend to implement.
-Test Last testing is the practice of writing tests after code has been written to check the functionality of the written code.
-These tests are then used to check the written code for any potential errors which are then fixed by the developer until no errors are found by the tests.
-Test last testing is considered simple to implement and can allow the programmer to focus on writing tests based on areas of his code that he may be concerned about.
-
-\subsection{Agile Oriented Development and Testing}
-
-In the late 1990s some software developers started to criticize the sequence-oriented waterfall model, complaining that it was too brittle and inflexible to meet the demands of the standard customer.
-In response to these critiques of the waterfall model a new model for developing software emerged, the agile development model.
-This new development model, based on the tenets of the agile manifesto \mycomment{cite manifesto here!}, promoted the idea that all actions of development should not occur in an ordered sequence but in a series of time-boxed iterations where in each iteration setting requirements, product design, product implementation and product testing would occur simultaneously.
-The goal of each iteration is to produce a demonstrable working product to show a customer and receive feedback on the product.
+In the late 1990s, a group of software developers started to criticize the phase-oriented waterfall model, complaining that it was too brittle and inflexible to meet the demands of the most customers.
+In response to these critiques of the waterfall model, a new model for developing software emerged, the agile development model.
+This new development model, based on the tenets of the agile manifesto ~\cite{agile:xxx}, promoted the idea that all actions of development should not occur within an ordered sequence of phases, but instead a series of time-boxed iterations where, in each iteration, developers set requirements, design a product, make the product, and test the product based on feedback from the previous iteration.
+The goal of each iteration is to produce a demonstrable sub-product to show a customer and to receive feedback on that sub-product.
 Some current development practices that are considered agile are Extreme Programming and Scrum.
@@
 Due to the changes in Agile programming, test-last testing was pushed aside in favor of a different style known as test-first testing.
-In 1996 \mycomment{might be 2001} with the release of the agile development practice Extreme Programming, the idea of test first testing, implemented in test driven development, started to become popular for the first time ~\cite{Hammond:2012}.
+In 2001, with the release of the agile development prachttp://start.fedoraproject.org/tice Extreme Programming, the idea of test-first testing, implemented in test driven development, started to become popular for the first time ~\cite{Hammond:2012}.
 Test-first testing is the practice of writing tests before code has been written and then writing code to make the tests pass.
-It should be noted that since tests occur before development test-first testing tends to be heavily linked to development methods and thus the most common test-first models also include development elements as well.
+It should be noted that since tests are written before production code, test-first testing tends to be heavily linked to development methods and thus the most common test-first models also include development elements as well.
 
-The most well known and most used test-first model is that of test driven development or TDD for short.
-In its current use in the field, the phrase TDD is mostly used as a blanket term for any sort of development practice that uses test-first testing to drive code development.
-That being said it is important to note that although TDD has been turned into a vague term in this paper we will focus more on its original specific definition.
-\mycomment{add more once we get to section 3}
+The most well known and used test-first model is that of test driven development, or TDD for short.
+In the original TDD methodology, the developer uses a series of steps to develop his code.
+The series begins once all current tests pass, or a new project is started.
+When this occurs, the programmer writes a new failing test that tests the simplest functionality the programmer wishes to add to their code.
+Once the test has been written and the test fails, the programmer then writes the minimal amount of code to make the test pass.
+After this step the programmer streamlines his solution and integrates it with other parts of his code.
+This series of steps repeats until the code is complete ~\cite{Hammond:2012}.
+In its current use in the field, TDD no longer has the uniform meaning that is described above.
+Instead it now refers to a loose collection of practices that roughly follow some or all of the guidelines given above.
+This means that although there is a specific definition for TDD, we can not assume that a study used the formal definition of TDD unless the TDD process is outlined within the study.
 
 \section{Research Data on Test-First vs Test-Last Testing}
-In this section we attempt to attempt to obtain useful conclusions between test-first and test-last testing by observing current research articles.
-In order to accomplish this we will first start with a data subsection that will go over the three main studies that will be used to draw our conclusions.
-Next we will have an analysis subsection that will summarize the results of the study in terms of our three main empirical measurements: code coverage, code correctness, and total development time.
-We will then have a discussion subsection that will take about potential issues with summarizing the data given in the research articles and then we shall proceed with a conclusion subsection where we will talk about potential conclusions we can draw from the data.
+In this section we will attempt to obtain useful comparisons between test-first and test-last testing by reviewing current research articles.
+We will first review three main studies.
+Next, we will summarize the results of our studies in terms of our three main empirical measurements: code coverage, code correctness, and total development time.
+Afterwards we will discuss the potential issues with summarizing the data given in the research articles and then we will talk about potential useful comparisons we can draw from the data.
 
 \subsection{Data}
- In 2012 Lemos et.
-al ~\cite{Lemos:2012} conducted a study on sixth semester students to see if test-first testing would increase code coverage, increase testing code size, increase code correctness, and/or decrease in total development time in auxiliary functions (functions with 10-200 lines of code) compared to test-Last testing.
-In this study 39 sixth semester students, knowledgeable in testing techniques and participants of two 100 minute long test first modules, were asked to complete coding challenges over two sessions.
-In the first session, half the students implemented testing-first testing while the other half implemented test-last testing to help solve the problem.
-In the second section the students were asked to switch roles and given a different coding challenge.
-In order to reduce domain knowledge bias amongst the coding challenges three different challenge question types were handed out at each session.
-The results of the study, which were analyzed for statistical significance using the Wilcoxon/Mann-Whitney non-parametric signed-rank paired test, found that test-first testing significantly increased code coverage, testing code size and code development time with no significant increase in code correctness.
 
-In 2010 Kollanus ~\cite{Kollanus:2010} conducted a literary review of TDD.
-Using standard literary review methods, Kollanus found and reviewed forty different experiments in scientific journals, magazines, and conference proceedings that provided empirical evidence for TDD.
-In the review Kollanus found that the most used measurements in these experiments could be clumped into three measurements: external code quality, internal code quality, and productivity.
-In the study Kollanus observed that there was weak support for improved external code quality, very weak support for internal quality, and support for an increased amount of development time.
+This subsection contains summaries of important studies that will be used later in the paper to draw out important conclusions and comparisons between test-first and test-last testing.
+During this section we will make multiple references to test last development methods which we will refer to as TLD.
 
-In 2012 Hellmann et.
+\subsubsection{Review by Kollanus}
+
+First, is a 2010 review by Kollanus \cite{Kollanus:2010}.
+In this review, Kollanus,  reviewed forty different experiments in scientific journals, magazines, and conference proceedings that provided empirical evidence comparing TDD  to TLD.
+Each study was assigned to one of three categories: Controlled experiments, case studies, and others.
+Controlled experiment articles were articles that went over a conducted controlled experiment.
+Case study articles were articles that summarized received data from a group or situation over a period of time.
+Other articles were any articles that did not fit into the above categories and were either non-controlled experiments or surveys.
+Overall there were 14 controlled experiments, 14 case studies, and 12 studies defined as other.
+
+In this review, Kollanus focuses on three different code quality measurements: external code quality, productivity, and internal code quality.
+External code quality, as defined by Kollanus, refers to how many errors are found in participants code.
+This definition is the same as the definition for code correctness.
+Because of this, we can use External code quality as measurement of code correctness.
+Productivity, as defined by Kollanus, measures multiple properties related to how efficient the product code was to create.
+This includes the measurement of total development time.
+Although productivity is not strictly total development time, Kollanus wrote her conclusions in productivity based mainly on development time which suggests that the majority of productivity studies focused on total development time.
+This means that the data given could be useful as a rough estimate of the effectiveness of total development time.
+Internal code quality describes a wide set of measurements that measure code quality from a testing and development standpoint.
+Code coverage is one of these measurements but is only mentioned briefly in a summary of one article.
+Because of this, we are not able to assume that internal quality is a good estimate for code coverage.
+Thus, we will be ignoring the overall results of internal code quality and focus on the summary of the one study.
+
+ In the study, Kollanus concluded that there was weak support for improved external code quality in TDD methods compared to TLD methods.
+This conclusion was based on the fact that out of the 22 studies that focused on external code quality, only 6 studies concluded that TDD did not increase external code quality.
+However Kollanus is weary about this conclusion because out of the 7 controlled experiments that considered external code quality only 2 of them report an increase in code quality.
+This was concerning to Kollanus because the data from the controlled experiment articles are generally more accurate than the other two types of articles.
+
+In terms of productivity, Kollanas suggests that TDD \textit{may} be less productive.
+This conclusion was drawn from 23 studies where 11 of the studies claimed that TDD decreases productivity, 7 of the studies say that there was no difference in productivity, and 5 studies that say there was an increase of productivity.
+In this case the controlled experiment data accurately reflects these numbers as out of the 10 controlled experiments 2 of them claimed increased productivity, 4 of them claim no difference, and 4 of them claim decreased productivity.
+It is interesting to note that although decreased productivity was the most common result in the review, the majority of the studies in the review state that TDD does not decrease productivity.
+
+There was one study that Kollanus mentions that considered code coverage.  This mention occurs in the internal code quality section and noted that TDD improved test coverage.
+
+\subsubsection{Review by Jeffries and Mitnik}
+
+In 2012, Hellmann et.
 al conducted a summary on the current status of research of TDD.
-Within Hellmann’s background there is a mention of a summary study produced by Jeffries and Mitnik that found in general that TDD largely resulted in an increase in quality, but one study they identified showed instead that TDD resulted in a strong negative impact on quality.
-Additionality while they showed that TDD could reduce the amount of effort required by up to 27\%, most studies found an increase in effort of up to 100\%.
+Within Hellmann's background there is a mention of a summary study produced by Jeffries and Mitnik that found in general that TDD largely resulted in an increase in quality, but one study they identified showed instead that TDD resulted in a strong negative impact on quality.
+Additionally while they showed that TDD could reduce the amount of effort required by up to 27\%, most studies found an increase in effort of up to 100\%.
+
+\subsubsection{Experiment by Lemos et al}
+
+In 2012, Lemos et.
+al ~\cite{Lemos:2012} conducted a study on computer science students to see if test-first testing would significantly impact code coverage,  code correctness, and/or total development time in auxiliary functions (functions with 10-200 lines of code).
+This study used 39 third-year computer science students knowledgeable in testing techniques.
+Each student took part in  two 100 minute test-first modules and were then asked to complete coding challenges over two sessions.
+In the first session, half the students used test-first methods while the other half used test-last methods.
+In the second session the students were asked to switch roles and given a different coding challenge.
+
+Code coverage and total development time were measured as mentioned in Section II.
+Code correctness was measured by running a set acceptance tests designed by the experimenter against a participant's code.
+Each code entry was then given a score based on how many acceptance tests passed.
+The scores were either 0 (all test cases fail), 0.5 (some test cases fail) or 1 (all test cases pass).
+This style of measuring code correctness has been used in previous studies and is called the \textit{Functional Test Set Success Level} scale.
+
+It was found that code coverage on average was 40\% higher when test-first testing was used.
+This percentage increase was found to be statistically significant.
+Thus, it was concluded that test-first methods produce higher test coverage then test-last methods.
+On total development time, it was found that test-first code took 12\% longer to write then test-last code.
+This result was found to be significantly significant which lead to the conclusion that test-first methods took longer to implement then test-last methods.
+In terms of code correctness, the only difference found was that the test-last code had one more correct implementation of code then the test-first code.
+In the case of test last code correctness, two submissions scored a 0, thirty-three submissions scored a 0.5, and five submissions scored a 1.
+Test first code correctness scored slightly worse with two submissions that scored a 0, thirty-four submissions that scored a 0.5, and four submissions that scored a 1.
+These results were not found to be statistically significant and thus it was concluded that test first testing had no impact on code quality compared to test last testing.
+
+\subsubsection{Opinion Study by Janzen}
 
@@
 it was acknowledged that increased test coverage had occurred in test-first testing methods compared to their test-last counterparts.
-All three papers brought up up total development time and noted that test-first testing tended to increase development time, although both Kollanus’s study and Hellmann’s study provide notable contradictory studies to this conclusion.
+All three papers brought up total development time and noted that test-first testing tended to increase development time, although both Kollanus's review and Hellmann's review provide notable contradictory studies to this conclusion.
 All three studies in one way or another also brought up code correctness in the form of external code quality in Kollanus or in the form of code quality in the case of Hellmann et al.
 In this case the results were across the board where Lemos et al.
-stated that there was no code correctness difference between test first and test last whereas Kollanus said there was weak evidence that test-first testing produces more correct code then test-last testing.
+stated that there was no code correctness difference between test first and test last whereas Kollanus said there was weak evidence that test-first testing, in the form of TDD, produces more correct code then test-last testing.
 Hellmann et al provided contradictory evidence backing that in most studies test-first testing has shown better code correctness then test last testing but there was a noted study where test-first testing was less code correct then test-last testing.
@@
 \subsection{Discussion}
-Many summaries of TDD research ~\cite{Hammond:2012, Hellman:2012, Kettunen:2010} , including the literary review done by Kollanus ~\cite{Kollanus:2010}, have noted that one confounding factor to their ability to summarize the research in the given field of test-first testing has been that many studies are contradictory to one another.
-For example in ~\cite{Hellman:2012},  there is a mention of previous summary studies that have had previous issues with conflicting data.
+Many summaries of TDD research ~\cite{Hammond:2012, Hellman:2012, Kettunen:2010} , including the literary review done by Kollanus ~\cite{Kollanus:2010}, have noted that it is hard to draw conclusions with the given research because that many studies that compare test-first to test-last testing are contradictory to one another.
+For example in ~\cite{Hellman:2012},  there is a mention of previous summary studies that have had issues with conflicting data.
 This is then followed by a series of examples of data that has conflicted in other studies in the past.
-In Kollanus’s literature review ~\cite{Kollanus:2010} Kollanus points out that many articles in her study contradicted other articles which made it hard to make sound conclusions from the research given and may be a confounding factor in her study.
-In most current studies this seems to occur because of two main factors: participant experience and improper study implementation and documentation.
+In Kollanus's literature review ~\cite{Kollanus:2010}, Kollanus points out that many articles in her study contradicted other articles which made it hard to make sound conclusions from the research given and may be a confounding factor in her study.
+This seems to occur because of two main factors: participant experience and improper study implementation and documentation.
 
@@
 Most studies, claimed Kollanus, had maybe one or two lines describing their TDD methodology or just simply claimed that they used TDD.
-Considering that TDD can be implemented in many ways, as discussed earlier in the background, Kollanus states that the lack of documentation for TDD implementation means that, as a reviewer, she has no idea if the test methods used to produce the results were similar or different thus potentially confounding some of her conclusions.
+Considering that TDD can be implemented in many ways, as discussed earlier in the background, Kollanus states that the lack of documentation for TDD implementation means that, as a reviewer, she has no idea if the test methods used to produce the results were similar or different, potentially confounding some of her conclusions.
 Another issue similar to this is the conformance issue, brought up in (cite here).
@@
 
+\section{Test Driven Development}
+\mycomment{newly added section, still rough at best}
+
+Test-first testing in most of these studies are being done in various forms of TDD.
+Because of this, the data for test-first testing has the potential to reflect traits of TDD which are independent of any traits in test-first testing.
+Although all the results have a potential to contain this issue one particular result seems to suggest it may be a trait of TDD, not test-first testing.
+This trait is that Test-first testing is more difficult then test last testing.
+
+\subsection{Studies noting TDD difficulty}
+Hammond et.
+al states in his research summary paper on TDD that: "TDD remains deceptively simple to describe but deeply challenging to put into practice effectively" `\cite{Hammond:2012}.
+Hammond reaches this conclusion with the use of multiple studies that show implecations that TDD is complex.
+Some of the studies used to draw this conclusion are summarized below.
+
+In the study by Janzen 2007, there was an interesting relation between two different measurement aspects of the opinion study.
+It was noted by Janzen that in every experiment, slightly less people said they would be more likely to implement TDD than people who said that TDD was a superior method to test-last methods.
+This means that more people thought that TDD was the better method to implement than  people who said they would actually implement TDD.
+Janzen attributed this phenomena as a testament to the difficulty of TDD as he found in the comments section multiple survey participants mentioned that they felt that TDD was too difficult or too different from what they normally do.
+
+In a study done by George and Williams,  \mycomment{cite this article!} multiple programming professionals from John Deere, RoleModel Software, and Ericsson participated in an experiment comparing TDD to waterfall development.
+In this study a 9 question survey was given out asking programmers what they thought about TDD and what was difficult with TDD.
+In this survey 56\% of the professionals noted that they had difficulty adapting to the TDD mindset when participating the study.
+In addition 23\% of the participants noted that they felt that the lack of upfront design in TDD was more of a hindrance then a help.
+
+In an online survey done by Aniche and Gerosa ~\cite{Aniche:2010}, 218 TDD programmers of differing skill levels were surveyed about their TDD practices and mistakes.
+In the survey it was found that TDD was not easy to follow as about 25\% of the programmers admitted to frequently or always making mistakes in following the traditional steps of TDD.
+Two examples of these mistakes include: forgetting to clean up their code after a test passes and writing tests that are too complex for effective TDD.
+
+\subsection{Reasons for TDD Difficulty}
+The data above shows fairly strong evidence that TDD is at least somewhat difficult to implement but none of the data above shows that test-first testing is the problem.
+In the first study, Janzen reveals that TDD was difficult for some people to implement.
+In the second study, George and Williams reveal that 56\% of the participants found the TDD mindset hard to adapt while only 23\% of the participants found testing before designing as a problem, suggesting that something other then test-first testing caused difficulty for at least 33\% of the participants.
+Whereas the third study shows that programmers have some trouble implementing the steps of TDD, which has nothing to do with testing before designing.
+Since difficulty in using TDD does not seem to be directly linked with test-first testing, that suggests that there is different issue causing TDD to be difficult to implement.
+
+One potential issue with TDD that is brought up by many developers is that it fails to explain the best way to implement tests to test your codes wanted functionality.
+This is best represented in a quote from Dan North's article \mycomment{cite Introducing BDD here}: "While using and teaching agile practices like test-driven development (TDD) on projects in different environments, I kept coming across the same confusion and misunderstandings.
+Programmers wanted to know where to start, what to test and what not to test, how much to test in one go, what to call their tests, and how to understand why a test fails."  This issue as shown in Dan North's article may reveal why TDD is so difficult.
+If the issue for why TDD is so difficult is similar to something like the issue explained above then perhaps they may be a better implementation other than TDD that exists.
+
 \section{evolutions to TDD}
@@
 
-As can been seen in the above sections there are some issues in the current implementation of TDD testing.
-Due to these issues new methods are starting to appear in the agile community that contain some of TDD's main tenets but have shifted enough away from TDD at some fundamental level that they are starting to receive their own names and classifications.
-Although there are many of TDD spin-offs, for this paper we chose to focus on three more popular spin offs, acceptance test driven development, Behavior driven development, and Agile Specification Driven Development.
+As can been seen in the above section there is an issue where TDD can be difficult to implement correctly.
+Due to this issue, new methods are starting to appear in the agile community that contain some of TDD's main tenets but have shifted enough away from TDD at some fundamental level that they are starting to receive their own names and classifications.
+Although there are many of TDD spin-offs, for this paper we chose to focus on two more popular spin offs, acceptance test driven development and behavior driven development.
+http://start.fedoraproject.org/
+\mycomment{add lack of research disclaimer}
 
-\subsection{Test Driven Development}
-Test driven development, as mentioned earlier is currently the most widely used test first testing and design methodology.
-Although in the previous cases we have been using TDD as a vague term in this section we will focus on the the original TDD methodology introduced in Extreme Programming.
-In the original TDD methodology the developer repeats a step of processes to develop his code.
-The process begins once all current tests pass or the new code has just been started.
-When this occurs the programmer writes a new failing test that tests the simplest functionality the programmer wishes to add to their code.
-Once the test has been written and the test fails, the programmer then writes the minimal amount of code to make the test pass.
-After this step the programmer proceeds to refactor any clutter within his solution and then repeats the process until the code is complete.
-
-Although TDD is a very popular style of test first testing, it should be noted that a perceived weakness to the TDD methodology exists.
-The main issue many people have with test driven development is that it fails to suggest a style of testing and how to best define an independent functionality in ones code plan.
-In the study ~\cite{Kettunen:2010} (cite others here as needed) people noted that TDD was difficult because they had a hard time deciding how to implement their tests effectively.
-Another related issue noted in ~\cite{Kettunen:2010} was that some individuals found it hard to figure out what needed to specifically be tested thus some found they wrote an overabundance of tests causing the process to become inefficient and thus take more time.
-Others found that focusing on the simplest functions first left serious potential problems to the end of development causing a serious time crunch problem for product release.
-Due to these issues with defining functionality and other issues within TDD new methods for implementing test first testing are being designed and implemented in an attempt to solve these problems.
 \subsection{Acceptance Test Driven Development}
 
-Perhaps the closest spin off to TDD, acceptance test driven development (ATDD), follows many of the ideas and expectations that regular TDD does, like writing tests first and following and using tests to define the development code.
-Unlike TDD though ATDD believes that in order to produce higher quality code and better customer satisfaction customers should write or define "acceptance" tests that must be passed before the product is considered finished.
+Perhaps the closest spin off to TDD, acceptance test driven development (ATDD), follows many of the ideas and expectations that regular TDD does, like writing tests first and using tests to define the development code.
+Unlike TDD, though, ATDD believes that in order to produce higher quality code, reduce coding confusion, and promote better customer satisfaction, customers should write or define "acceptance" tests.
+These acceptance tests are tests that are required to pass before the customer will accept the product and usually define the core functionality of the product being developed.
 This part of ATDD is noticeably different from the TDD ideology because in the TDD ideology only the developer should writes the tests and the tests should be written from simplest to hardest.~\cite{Hammond:2012}
 
-Although some people claim ATDD is a step in the right direction for TDD developers their are also many people who have complaints about ATDD.
-One of the major complaints is that ATDD is an ideal practice at best because customers will not take the time to create useful acceptance tests.~\cite{Hammond:2012} Yet others state that ATDD is very hard for both customers and developers to implement correctly because their is no effective common language between the two factions.(site new IEEE article here)  Because of these varied complaints some of the promoters of ATDD have moved away from it to support a newer similar TDD spin off, Behavior Driven Development or BDD.
+The shift from having the developer write all the tests to the customer having to write some of tests is ideally considered a win win.
+The shift of the workload of determining main functionality from the developer to the customer significantly reduces the difficulty of the TDD practice to the developer as they are no longer concerned about writing and determining all the functionality for their code.
+This method is also supposed to benefit the customer as it allows the customer to get the product they desire.
+One of the major complaints of ATDD is that it is an ideal practice at best because customers will not take the time to create useful acceptance tests.~\cite{Hammond:2012} Another complaint of ATDD is that its very hard for  customers to write useful acceptance tests  because they view functionality differently then developers because their knowledge of code is limited.(site new IEEE article here)  Because of these varied complaints, some of the promoters of ATDD have moved away from it to support a newer similar TDD spin off, Behavior Driven Development.
 
 \subsection{Behavior Driven Development}
-BDD is a new spin off of TDD that focuses on how to correctly implement the fundamental usefulness of TDD.
-In BDD the most important thing that it argues is that TDD is unsuccessful because it is vague and it focuses more on "testing the code" then showing the codes behavior.
-Thus BDD argues that solving the problem is to stop thinking about code in terms of testing but to understand code in terms of behavior.
-Because of this shift in ideas BDD uses different testing tools, such as JBehave for java or Cucumber for Ruby \cite{Soeken:2012}, compared to TDD that breaks code testing down into standard verbal sentences instead of basic testing language.
-The argued pros for this type of language usage is that it actually reduces much of the difficulty of tdd by allowing us to define the purpose of our code in more natural terms thus removing much of the confusion about what is important to test in TDD and the other pro is that now it is easier to communicate goals and functionality of the code to non computer science majors (IE Managers, Customers, Sales Reps, etc.) which allows for higher customer satisfaction.
-\mycomment{this is just a blurb on BDD and what it is and what perceived advantages it contains, this paragraph will get reworked as it does not fit in the current paper well}
+\mycomment{somewhat stubbed section atm}
+Behavior Driven Development, or BDD, is a new spin off of TDD that focuses on how to correctly implement the fundamental usefulness of TDD.
+The main difference between BDD and TDD is that TDD focuses on "testing the code" while BDD focuses more on defining a codes behavior.
+In order to do this, BDD focuses on defining a codes wanted functionality in plain English and then converts that plain English into tests using various programming tools and methods \mycomment{cite dan North and hammond article here}.
+This differs from TDD, because in TDD, their is no focus on defining code functionality in terms of plain English or even focusing specifically on code functionality.
 
-\mycomment{need to do more research here}
-
-\subsection{Agile Specification-Driven Development}
-The new testing style Agile Specification-Driven Development is a hybrid of two older testing styles TDD and DbC (design by contract).
-
-Advantage to merging the DbC system with TDD is it gives the designer a better idea about what is important to test \mycomment{and other advantages should also crop up in research on DbC}
-
-\mycomment{need to do research here, specifically it would be useful to do some research on DbC}
-
-\subsection{section conclusion}
-\mycomment{the goal here is to have a discussion over which TDD spin off might be most promising and for what reasons}
+The argued pros for BDD is that it can bypass many of the difficulties of TDD by allowing us to define the purpose of our code in more natural terms.
+This style removes much of the confusion about what is important to test in TDD and the other pro is that now it is easier to communicate goals and functionality of the code to non computer science majors (IE Managers, Customers, Sales Reps, etc.) which allows for higher customer satisfaction.
 
@@
 
-\subsection{Discussion of Implications of Section 3}
+\subsection{Discussi or BDDon of Implications of Section 3}
 \mycomment{stubbed for now}

### 2013-11-23 Paper_Draft_Alt.tex
pushing stuff for funsies

--- 2013-11-09 Paper_Draft.tex
@@
 
-Although there are many testing methods that currently exist, two particular testing methods are currently very popular: test-last testing and test-first testing.
+Although there are many testing methods that  exist, two particular testing methods are currently very popular: test-last testing and test-first testing.
 Test-last testing, used mostly in process oriented or waterfall development, is a testing method where testing is done after the software code has been written to ensure that the code works correctly.
@@
 
-Currently, there is much debate in the testing community about whether or not test-first or test-last testing is superior.
+\mycomment{this transition needs to be rewritten}
+Recently, there has been much debate in the testing community about whether or not test-first or test-last testing is superior.
 The goal of this paper is to attempt to give an overview of the current state of this debate by analyzing current research data concerning the advantages and disadvantages of each of these testing methods.
@@
 
-The rest of the paper is divided into four sections.
+The paper is divided into four sections.
 In section two of the paper we discuss what software testing is and current software development models with their supporting testing methods: test-first and test-last testing.
@@
 \subsection{Waterfall Development and Testing}
-
+\mycomment{present tense this paragraph}
 From as early as the 1970s, a popular software development model was the waterfall model, where software was developed in a series of phases.
 The waterfall development model is used because it is simple to implement correctly and is also considered time efficient because completing previous phases helps later phases become more efficient.
-The first phase in the standard waterfall model is the requirements phase where requirements are set by the customer or design company.
+The first phase in the standard waterfall model is the requirements phase, where requirements are set by the customer or design company.
 Next the design phase occurs in which the product is designed and then the product is built in the implementation phase.
@@
 
-The most well known and most used test-first model is that of test driven development or TDD for short.
+The most well known and most used test-first model is that of test driven development, or TDD for short.
 In its current use in the field, the phrase TDD is mostly used as a blanket term for any sort of development practice that uses test-first testing to drive code development.
-That being said it is important to note that although TDD has been turned into a vague term in this paper we will focus more on its original specific definition.
-\mycomment{This paragraph will need to be reworked once the TDD section is reworked, it is currently out of date.}
+That being said it is important to note that although TDD has been turned into a vague term, in this paper we will focus more on its original, specific definition.
+\mycomment{merge some of the ideas of this paragraph with the intro to the paragraph}
+
+Test driven development, as mentioned earlier is currently the most widely used test first testing and design methodology.
+Although in the previous cases we have been using TDD as a vague term in this section we will focus on the the original TDD methodology introduced in Extreme Programming.
+In the original TDD methodology, the developer repeats a step of processes to develop his code.
+The process begins once all current tests pass, or the new code has just been started.
+When this occurs, the programmer writes a new failing test that tests the simplest functionality the programmer wishes to add to their code.
+Once the test has been written and the test fails, the programmer then writes the minimal amount of code to make the test pass.
+After this step the programmer streamlines his solution with the other parts of his code.
+This series of steps repeats until the code is complete.
 
@@
 \subsection{Data}
-\mycomment{Reworked the paragraphs on the Kollanus study and the Lemos study in an attempt to better represent data in my paper, also planning on writting paragraphs for two other studies.
-Considering making this its own section due to its increase in size.}
 
@@
 In the review Kollanus focuses on three different measurements: external code quality, productivity, and internal code quality.
-In the article, Kollanus describes external code quality as the measurement of the number of passed researcher acceptance tests in a test set when run against code written by participants.
+In the article, Kollanus describes external code quality ahttp://start.fedoraproject.org/s the measurement of the number of passed researcher acceptance tests in a test set when run against code written by participants.
 Since this definition is equivalent to our definition for code-correctness this data can be used for code-correctness comparisons.
@@
 This conclusion was based on consistent evidence found in case studies that reported better external code quality in TDD methods.
-However, the support was considered weak because there were contradictory results found in some controlled experiments that recorded external code quality.
+However, the support was considered weak because there were contradictory results found in some higher quality controlled experiments that recorded external code quality.
 Productivity, as defined by Kollanus, is a measurement that measures how much code one writes in comparison to time spent.
@@
 Two of the three studies covered the topic of code coverage.
-In both studies by Kollanus and Lemos et al.
+In both shttp://start.fedoraproject.org/tudies by Kollanus and Lemos et al.
 it was acknowledged that increased test coverage had occurred in test-first testing methods compared to their test-last counterparts.
-All three papers brought up up total development time and noted that test-first testing tended to increase development time, although both Kollanus’s study and Hellmann’s study provide notable contradictory studies to this conclusion.
+All three papers brought up total development time and noted that test-first testing tended to increase development time, although both Kollanus’s study and Hellmann’s study provide notable contradictory studies to this conclusion.
 All three studies in one way or another also brought up code correctness in the form of external code quality in Kollanus or in the form of code quality in the case of Hellmann et al.
 In this case the results were across the board where Lemos et al.
-stated that there was no code correctness difference between test first and test last whereas Kollanus said there was weak evidence that test-first testing produces more correct code then test-last testing.
+stated that there was no code correctness difference between test first and test last whereas Kollanus said there was weak evidence that test-first testing, in the form of TDD, produces more correct code then test-last testing.
 Hellmann et al provided contradictory evidence backing that in most studies test-first testing has shown better code correctness then test last testing but there was a noted study where test-first testing was less code correct then test-last testing.
@@
 \subsection{Discussion}
-Many summaries of TDD research ~\cite{Hammond:2012, Hellman:2012, Kettunen:2010} , including the literary review done by Kollanus ~\cite{Kollanus:2010}, have noted that one confounding factor to their ability to summarize the research in the given field of test-first testing has been that many studies are contradictory to one another.
-For example in ~\cite{Hellman:2012},  there is a mention of previous summary studies that have had previous issues with conflicting data.
+Many summaries of TDD research ~\cite{Hammond:2012, Hellman:2012, Kettunen:2010} , including the literary review done by Kollanus ~\cite{Kollanus:2010}, have noted that it is hard to draw conclusions with the given research because that many studies that compare test-first to test-last testing are contradictory to one another.
+For example in ~\cite{Hellman:2012},  there is a mention of previous summary studies that have had issues with conflicting data.
 This is then followed by a series of examples of data that has conflicted in other studies in the past.
-In Kollanus’s literature review ~\cite{Kollanus:2010} Kollanus points out that many articles in her study contradicted other articles which made it hard to make sound conclusions from the research given and may be a confounding factor in her study.
-In most current studies this seems to occur because of two main factors: participant experience and improper study implementation and documentation.
-
-In the case of participant experience there seems to be a skill gap between test first and test last testing.
-In an opinion study by ~\cite{Janzen:2007} it was found that a majority college students said they were not comfortable with and had trouble with comprehending TDD, even after doing a coding exercise with it.
-The majority of students also said that they were much more comfortable implementing test last methods and would prefer to use test last methods in the future.
-Coding professionals, on the other hand, actually preferred test first programming and were much more comfortable with it than students.
-They also stated they felt it was relatively easy to understand while students did not.
-In another opinion study ~\cite{Kettunen:2010}, by Kettunen et al, a common theme throughout the paper was examples of individual testimonies and research conclusions that pointed to test first testing being harder to implement than test last testing.
-Considering that most conducted experiments focus on college students and industry professionals, who have vast experience differences, the fact that test first testing may be harder to implement for college students compared to industry professionals may create differing results between studies.
-This idea is considered in ~\cite{Hammond:2012} and ~\cite{Kettunen:2010}, but no summary papers where found considering this result and comparing student participant experiment results to professional participant experiment results.
+In Kollanus’s literature review ~\cite{Kollanus:2010}, Kollanus points out that many articles in her study contradicted other articles which made it hard to make sound conclusions from the research given and may be a confounding factor in her study.
+This seems to occur because of two main factors: participant experience and improper study implementation and documentation.
 
@@
 
-\subsection{Conclusion}
-
-Due to current issues in contradictory data it is hard to make solid conclusions about the advantages and disadvantages of using test first testing instead of test last testing.
-What can be done instead is we can consider certain trends in research that can suggest potential answers.
-Most studies and summaries seem to agree that test-first testing tends to produce more test coverage compared to test-last testing and that test first testing tends to take longer to write tests compared to test first testing.
-Two of the studies we looked at, the experiment by lemos and the literary review by Kollanus, point out these two trends.
-Also one or both trends have been noted through multiple papers ~\cite{Kettunen:2010, Hammond:2012, Hellman:2012} with no major contradictions found.
-Another convincing trend that has occurred is that in general test first testing seems to be harder to implement than test last testing as mentioned earlier.
-Although these trends seem to be clear, the rest seem to be more muddled.
-Code correctness for example seems to have no clear trend as multiple studies have reached contradictory conclusions to other studies.
-Another hot button topic that fails to show a trend one way or the other is the amount of time spent simply developing code as this again seems to have results across the board.
-From these trends we can consider that test first testing overall probably takes more time to write, increases code coverage, and is harder to implement than test-last testing and that other advantages or detriments may exist but they are unclear in current research.
+In the case of participant experience there seems to be a skill gap between test first and test last testing.
+In an opinion study by ~\cite{Janzen:2007} it was found that a majority college students said they were not comfortable with and had trouble with comprehending TDD, even after doing a coding exercise with it.
+The majority of students also said that they were much more comfortable implementing test last methods and would prefer to use test last methods in the future.
+Coding professionals, on the other hand, actually preferred test first programming and were much more comfortable with it than students.
+They also stated they felt it was relatively easy to understand while students did not.
+In another opinion study ~\cite{Kettunen:2010}, by Kettunen et al, a common theme throughout the paper was examples of individual testimonies and research conclusions that pointed to test first testing being harder to implement than test last testing.
+Considering that most conducted experiments focus on college students and industry professionals, who have vast experience differences, the fact that test first testing may be harder to implement for college students compared to industry professionals may create differing results between studies.
+This idea is considered in ~\cite{Hammond:2012} and ~\cite{Kettunen:2010}, but no summary papers where found considering this result and comparing student participant experiment results to professional participant experiment results.
 
 \section{Test Driven Development}
+Test-first testing in most of these studies are being done in various forms of TDD.
+Because of this, the data for test-first testing is biased toward any strengths or weaknesses that TDD contains.
+Although there are many small strengths and weaknesses in using TDD, one particular weakness of TDD has become noted directly or indirectly now in multiple studies.
+That weakness is best described in Hammond et.
+al which states that: "TDD remains deceptively simple to describe but deeply challenging to put into practice effectively" `\cite{Hammond:2012}.
 
-\mycomment{currently this section is more of a stub then anything.
-This section should be fleshed out over the weekend.
-Although I am not sure on formatting or precise content yet, many articles have noted that TDD takes longer and is harder to implement because it is unhelpful in determining what parts of the code/problem need to be tested and what is the best way of implementing those tests.}
-Test driven development, as mentioned earlier is currently the most widely used test first testing and design methodology.
-Although in the previous cases we have been using TDD as a vague term in this section we will focus on the the original TDD methodology introduced in Extreme Programming.
-In the original TDD methodology the developer repeats a step of processes to develop his code.
-The process begins once all current tests pass or the new code has just been started.
-When this occurs the programmer writes a new failing test that tests the simplest functionality the programmer wishes to add to their code.
-Once the test has been written and the test fails, the programmer then writes the minimal amount of code to make the test pass.
-After this step the programmer streamlines his solution with the other parts of his code.
-This series of steps repeats until the code is complete.
+\subsection{Studies noting TDD difficulty}
+\mycomment{note this study either needs to be previously discussed or this paragraph needs to be expanded to include the study}
+The fact that TDD is challenging to use effectively can be seen in multiple studies.
 
-~\mycomment{this next paragraph is an extremely rough draft that was written before this section became a section and currently makes little to no sense, comments are not needed for this paragraph}
+In the study by Janzen 2007, there was an interesting relation between two different measurement aspects of the opinion study.
+It was noted by Janzen that in every experiment, slightly less people said they would be more likely to implement TDD than people who said that TDD was a superior method to test-last methods.
+This means that more people thought that TDD was the better method to implement than  people who said they would actually implement TDD.
+Janzen attributed this phenomena as a testament to the difficulty of TDD as he found in the comments section multiple survey participants mentioned that they felt that TDD was too difficult or too different from what they normally do.
 
-Although TDD is a very popular style of test first testing, it should be noted that a perceived weakness to the TDD methodology exists.
-The main issue many people have with test driven development is that it fails to suggest a style of testing and how to best define an independent functionality in ones code plan.
-In the study ~\cite{Kettunen:2010} (cite others here as needed) people noted that TDD was difficult because they had a hard time deciding how to implement their tests effectively.
-Another related issue noted in ~\cite{Kettunen:2010} was that some individuals found it hard to figure out what needed to specifically be tested thus some found they wrote an overabundance of tests causing the process to become inefficient and thus take more time.
-Others found that focusing on the simplest functions first left serious potential problems to the end of development causing a serious time crunch problem for product release.
-Due to these issues with defining functionality and other issues within TDD new methods for implementing test first testing are being designed and implemented in an attempt to solve these problems.
+In a study done by George and Williams,  \mycomment{cite this article!} multiple programming professionals from John Deere, RoleModel Software, and Ericsson participated in an experiment comparing TDD to waterfall development.
+In this study a 9 question survey was given out asking programmers what they thought about TDD and what was difficult with TDD.
+In this survey 56\% of the professionals noted that they had difficulty adapting to the TDD mindset when participating the study.
+In addition 23\% of the participants noted that they felt that the lack of upfront design in TDD was more of a hindrance then a help.
+
+\mycomment{rewrite this after reviewing article, the summary given by hammond kinda sucks}
+In an online survey done by Aniche and Gerosa ~\cite{Aniche:2010}, 218 TDD programmers of differing skill levels were surveyed about their TDD practices and mistakes.
+In the survey it was found that TDD was not easy to follow as about 25\% of the programmers admitted to frequently or always making mistakes in following the traditional steps of TDD.
+Two examples of these mistakes include: forgetting to clean up their code after a test passes and writing tests that are too complex for effective TDD.
+
+\mycomment{A conclusion paragraph would be nice here}
+\subsection{Reasons for TDD Difficulty}
+The main problem that makes TDD so difficult to implement correctly is that it fails to explain the best way to implement tests to test your codes wanted functionality.
+This is best represented in a quote from Dan North's article \mycomment{cite Introducing BDD here}.
+"While using and teaching agile practices like test-driven development (TDD) on projects in different environments, I kept coming across the same confusion and misunderstandings.
+Programmers wanted to know where to start, what to test and what not to test, how much to test in one go, what to call their tests, and how to understand why a test fails."  Another quote that reveals this difficulty can be found in ~\cite{Hammond:2012}.
+"Design coordinates interacting entities.
+Choosing test cases that will adequately fill this role is difficult.
+It is true that a test case is an unambiguous requirement, but is it the correct requirement?"  This difficulty causes many problems when people attempt to implement TDD and potentially makes test first testing harder to implement correctly then it needs to be.
 
@@
 
-As can been seen in the above sections there are some issues in the current implementation of TDD testing.
-Due to these issues new methods are starting to appear in the agile community that contain some of TDD's main tenets but have shifted enough away from TDD at some fundamental level that they are starting to receive their own names and classifications.
-Although there are many of TDD spin-offs, for this paper we chose to focus on three more popular spin offs, acceptance test driven development, Behavior driven development, and Agile Specification Driven Development.
+As can been seen in the above section there is an issue where TDD can be difficult to implement correctly.
+Due to this issue, new methods are starting to appear in the agile community that contain some of TDD's main tenets but have shifted enough away from TDD at some fundamental level that they are starting to receive their own names and classifications.
+Although there are many of TDD spin-offs, for this paper we chose to focus on two more popular spin offs, acceptance test driven development and behavior driven development.
 
@@
 
-\mycomment{need to do more research here}
+\section{Conclusion}
 
-\subsection{Agile Specification-Driven Development}
-The new testing style Agile Specification-Driven Development is a hybrid of two older testing styles TDD and DbC (design by contract).
-
-Advantage to merging the DbC system with TDD is it gives the designer a better idea about what is important to test \mycomment{and other advantages should also crop up in research on DbC}
-
-\mycomment{need to do research here, specifically it would be useful to do some research on DbC}
-
-\subsection{section conclusion}
-\mycomment{the goal here is to have a discussion over which TDD spin off might be most promising and for what reasons}
+Due to current issues in contradictory data it is hard to make solid conclusions about the advantages and disadvantages of using test first testing instead of test last testing.
+What can be done instead is we can consider certain trends in research that can suggest potential answers.
+Most studies and summaries seem to agree that test-first testing tends to produce more test coverage compared to test-last testing and that test first testing tends to take longer to write tests compared to test first testing.
+Two of the studies we looked at, the experiment by lemos and the literary review by Kollanus, point out these two trends.
+Also one or both trends have been noted through multiple papers ~\cite{Kettunen:2010, Hammond:2012, Hellman:2012} with no major contradictions found.
+Another convincing trend that has occurred is that in general test first testing seems to be harder to implement than test last testing as mentioned earlier.
+Although these trends seem to be clear, the rest seem to be more muddled.
+Code correctness for example seems to have no clear trend as multiple studies have reached contradictory conclusions to other studies.
+Another hot button topic that fails to show a trend one way or the other is the amount of time spent simply developing code as this again seems to have results across the board.
+From these trends we can consider that test first testing overall probably takes more time to write, increases code coverage, and is harder to implement than test-last testing and that other advantages or detriments may exist but they are unclear in current research.
 
@@
 \subsection{Discussion of Implications of Section 3}
+
 \mycomment{stubbed for now}

### 2013-11-24 Paper_Draft.tex
should be ready to go.

@@
 
+Last, is a 2007 opinion study by Janzen \cite{Janzen:2007}.
+In this opinion study, Janzen polled participants from six experiments to determine whether they preferred testing with test-first methods or test-last methods.
+Five of these experiments were conducted on students at the University of Kansas while one of the experiments was conducted on professional programmers from a Fortune 500 company.
+Each participant was asked six questions where they had to choose between test-first and test-last testing.
+The questions were:
+
+\begin{itemize}
+\item which approach they would choose in the future (\textbf{Choice})
+\item which approach was the best for the project(s) they
+completed (\textbf{BestApproach})
+\item which approach would cause them to more thoroughly
+test a program (\textbf{ThoroughTesting})
+\item which approach produces a correct solution in less
+time (\textbf{Correct})
+\item which approach produces code that is simpler, more
+reusable, and more maintainable (\textbf{Simpler})
+\item which approach produces code with fewer defects
+
+(\textbf{FewerDefects})
+\end{itemize}
+
+The results of the study are shown in figure 1.
+The results were divided between a beginner group (students within their first few code classes) and a mature group (students nearing graduation and professionals).
+In the results section Janzen concluded that beginner coding students have a preference toward test-last testing while more mature coding students and industry professionals show a preference toward test-first testing.
+
 \subsection{Analysis}

### 2013-11-24 Paper_Draft.tex
Finished rewriting data section and first paragraph of discussion.

@@
 \subsection{Waterfall Development and Testing}
-\mycomment{present tense this paragraph}
+
 A popular software development model that was developed in the 1970's is the waterfall software development model, where software is developed in a series of phases.
@@
 Due to the fact that these processes often put testing at the end of development, a certain type of testing method, test-last testing, was the only testing method that made sense to use.
-http://start.fedoraproject.org/
+
 Test-last testing is currently a popular testing method and is usually the first testing method that people tend to implement.
@@
 
-\subsubsection{Review by Jeffries and Mitnik}
+%\subsubsection{Review by Jeffries and Mitnik}
 
-In 2012, Hellmann et.
+%In 2012, Hellmann et.
 al conducted a summary on the current status of research of TDD.
@@
 
+\begin{figure}
+\centering
+\psfig{file=janzengraph1.pdf,width =3in}
+\caption{Results for beginner programmers in Janzen's study}
+\end{figure}
+
+\begin{figure}
+\centering
+\psfig{file=janzengraph2.pdf,width =3in}
+\caption{Results for Mature programmers in Janzens study}
+\end{figure}
+
 \subsubsection{Opinion Study by Janzen}
@@
 \item which approach produces a correct solution in less
-time (\textbf{Correct})
+time (\textbf{Correctness})
 \item which approach produces code that is simpler, more
@@
 
-The results of the study are shown in figure 1.
-The results were divided between a beginner group (students within their first few code classes) and a mature group (students nearing graduation and professionals).
-In the results section Janzen concluded that beginner coding students have a preference toward test-last testing while more mature coding students and industry professionals show a preference toward test-first testing.
+The results were divided between a beginner group (students within their first few code classes) and a mature group (students nearing undergraduate graduation and professionals).
+The results for the beginner group can be seen in figure 1 and the results for the mature group can be seen in figure 2.
+Each graph shows the distribution of all of the participants answers to the six questions where each question is referenced by its bold keyword.
+With the data given in the figures, Janzen was able to conclude that beginner coding students have a preference toward test-last testing while more mature coding students and industry professionals show no significant preference toward either testing method.
 
-\subsection{Analysis}
-Two of the three studies covered the topic of code coverage.
+%\subsection{Analysis}
+%Two of the three studies covered the topic of code coverage.
 In both studies by Kollanus and Lemos et al.
@@
 
-\subsection{Discussion}
-Many summaries of TDD research ~\cite{Hammond:2012, Hellman:2012, Kettunen:2010} , including the literary review done by Kollanus ~\cite{Kollanus:2010}, have noted that it is hard to draw conclusions with the given research because that many studies that compare test-first to test-last testing are contradictory to one another.
-For example in ~\cite{Hellman:2012},  there is a mention of previous summary studies that have had issues with conflicting data.
-This is then followed by a series of examples of data that has conflicted in other studies in the past.
-In Kollanus's literature review ~\cite{Kollanus:2010}, Kollanus points out that many articles in her study contradicted other articles which made it hard to make sound conclusions from the research given and may be a confounding factor in her study.
-This seems to occur because of two main factors: participant experience and improper study implementation and documentation.
+\subsection{Understanding Contradictions}
+Many summaries of TDD research ~\cite{Hammond:2012, Hellman:2012, Kettunen:2010, Kollanus:2010} have noted that studies that compare test-first to test-last testing tend to contradict one another.
+Kollanus points out that articles in her review contradicted other articles which made it hard to make sound conclusions from the research given, especially on the subject of productivity.
+The amount of contradictions concerned Kollanus enough that she noted the amount of contradictions as a potential confounding factor for her results.
+Contradictions also existed between our review by Kollanus and our experiment by Lemos et al.
+Although both articles seem to agree that code coverage is increased with test-first methods neither study seems to totally agree on the attributes of code correctness or total development time.
+Since contradictions occur often in the results of our study we considered it to be important to explore some current theories on why these contradictions exist.
+In particular we will be exploring two potential causes: participant experience and improper study implementation/ documentation.
 
 In the case of participant experience there seems to be a skill gap between test first and test last testing.
-In an opinion study by ~\cite{Janzen:2007} it was found that a majority college students said they were not comfortable with and had trouble with comprehending TDD, even after doing a coding exercise with it.
-The majority of students also said that they were much more comfortable implementing test last methods and would prefer to use test last methods in the future.
-Coding professionals, on the other hand, actually preferred test first programming and were much more comfortable with it than students.
+This gap is best shown by the results from Janzen  Coding professionals, on the other hand, actually preferred test first programming and were much more comfortable with it than students.
 They also stated they felt it was relatively easy to understand while students did not.

### 2013-11-26 Paper_Draft.tex
Final Draft for monday.

@@
 
-\keywords{ACM proceedings, \LaTeX, text tagging}
+\keywords{Test Driven Development, Behavior Driven Development, Test-First, Test-Last, Agile, Waterfall, Testing, \LaTeX, text tagging}
 
@@
 
-The paper is divided into five sections.
+The paper is divided into four sections.
 In Section II we discuss what software testing is and current software development models with their supporting testing methods: test-first and test-last testing.
 In Section III we will provide an analysis of the data explaining the potential advantages and disadvantages of test-first testing compared to test-last testing.
-In Section IV we discuss issues of using test driven development to implement test-first testing.
-In Section V we will go over new testing methodologies based on test-first development.
-In Section VI we will provide conclusions and suggestions for further research in the field.
+In Section IV we discuss issues of using test driven development to implement test-first testing and will go over a new test-first methodology called behavior driven development.
+In Section V we will provide conclusions and suggestions for further research in the field.
 
@@
 Due to the changes in Agile programming, test-last testing was pushed aside in favor of a different style known as test-first testing.
-In 2001, with the release of the agile development prachttp://start.fedoraproject.org/tice Extreme Programming, the idea of test-first testing, implemented in test driven development, started to become popular for the first time ~\cite{Hammond:2012}.
+In 2001, with the release of the agile development practice Extreme Programming, the idea of test-first testing, implemented in test driven development, started to become popular for the first time ~\cite{Hammond:2012}.
 Test-first testing is the practice of writing tests before code has been written and then writing code to make the tests pass.
@@
 Since contradictions occur often in the results of our study we considered it to be important to explore some current theories on why these contradictions exist.
-In particular we will be exploring two potential causes: participant experience and improper study implementation/ documentation.
+In particular we will be exploring two potential causes: method difficulty and improper study implementation/ documentation.
 
-In the case of participant experience there seems to be a skill gap between test first and test last testing.
-This gap is best shown by the results from Janzen  Coding professionals, on the other hand, actually preferred test first programming and were much more comfortable with it than students.
-They also stated they felt it was relatively easy to understand while students did not.
-In another opinion study ~\cite{Kettunen:2010}, by Kettunen et al, a common theme throughout the paper was examples of individual testimonies and research conclusions that pointed to test first testing being harder to implement than test last testing.
-Considering that most conducted experiments focus on college students and industry professionals, who have vast experience differences, the fact that test first testing may be harder to implement for college students compared to industry professionals may create differing results between studies.
-This idea is considered in ~\cite{Hammond:2012} and ~\cite{Kettunen:2010}, but no summary papers where found considering this result and comparing student participant experiment results to professional participant experiment results.
+Method difficulty is considered a potential confounding factor when comparing test-first and test-last testing ~\cite{Hammond:2012, Kettunen:2010}.
+In particular, there is concern that test-first testing may be more difficult for beginning coders than test-last testing.
+This is a concern because most study data comes from two sources with very different skill levels: industry professionals and college students.
+In the study by Janzen et al, figures 1 and 2 show very different results between beginning and mature programmers.
+In particular, beginner programmers did not report a very positive experience with test-first testing compared mature programmers.
+This suggests that beginning coders may not be getting all the benefits from test-first methods that the more mature coders received.
+If this is the case, then any summary method with mixed results between beginning and advanced programmers has a potential to have contradictory results.
 
-The other problem that has plagued summary papers of test first testing is that most studies either lack documentation in key locations or do not effectively handle the participant conformance issue which comes up in TDD studies.
-In the literary review by Kollanus, she acknowledges a frustrating lack of information in many studies on how TDD was implemented.
-Most studies, claimed Kollanus, had maybe one or two lines describing their TDD methodology or just simply claimed that they used TDD.
-Considering that TDD can be implemented in many ways, as discussed earlier in the background, Kollanus states that the lack of documentation for TDD implementation means that, as a reviewer, she has no idea if the test methods used to produce the results were similar or different, potentially confounding some of her conclusions.
-Another issue similar to this is the conformance issue, brought up in (cite here).
-The conformance issue in TDD testing refers to the fact that few researchers make sure that TDD was implemented correctly or at all by its participants.
+The other problem that has plagued summary papers of test first testing is the lack documentation  and conformance of TDD methods in studies.
+Kollanus, in her review paper, acknowledges a frustrating lack of information in many studies on how TDD was implemented.
+Most articles, claimed Kollanus, had perhaps one or two lines describing their TDD methodology or only claimed that they used TDD.
+Considering that TDD can be implemented in many ways, this lack of documentation for TDD implementation creates a problem in accurately comparing two studies.
+This is an issue because multiple TDD methods may not preform equally, causing potentially contradicting results.
+This problem could be avoided if studies documented their TDD process.
+Another problem with TDD studies is that very few confirm their participants correctly implemented TDD~\cite{Hammond:2012}.
 For example, in the study by Lemos et.
-al the researchers acknowledged that one of their confounding factors was that they only asked students to write tests before code, thus a variety of different test first methods might have been used within the study making its data less conclusive.
-Overall the lack of documentation of how researchers specifically planned to implement TDD in their study and their lack of some sort of monitoring of whether the implementation was actually occurring greatly reduces the credibility of the research being done and allows for confounding factors to potentially occur.
+al, the researchers acknowledged that one of their confounding factors was that they only asked students to write tests before code, thus a variety of different test-first methods might have been used within the study making its data less conclusive.
+Overall the lack of documentation of how researchers specifically planned to implement TDD in their study and their lack of some sort of monitoring of whether the implementation was actually occurring greatly reduces the credibility of the experiment being done and allows for contradictory results to occur.
 
@@
 
-Due to current issues in contradictory data it is hard to make solid conclusions about the advantages and disadvantages of using test first testing instead of test last testing.
-What can be done instead is we can consider certain trends in research that can suggest potential answers.
-Most studies and summaries seem to agree that test-first testing tends to produce more test coverage compared to test-last testing and that test first testing tends to take longer to write tests compared to test first testing.
-Two of the studies we looked at, the experiment by lemos and the literary review by Kollanus, point out these two trends.
-Also one or both trends have been noted through multiple papers ~\cite{Kettunen:2010, Hammond:2012, Hellman:2012} with no major contradictions found.
-Another convincing trend that has occurred is that in general test first testing seems to be harder to implement than test last testing as mentioned earlier.
+Due to current issues in contradictory data it is hard to make solid conclusions about the advantages and disadvantages of using test-first testing instead of test-last testing.
+That being said current trends exist within the research that allow us to guess some of the traits of test-first methods compared to test-last methods.
+One result which was very clear in the research was that of code coverage.
+The results from Lemos et al and Kollanus suggest that test-first testing tends to produce more code coverage compared to test-last testing.
+This seems to be a fairly uncontested conclusion as~\cite{Kettunen:2010, Hammond:2012} reached the same conclusion and no contradictions were found in other articles.
+Another convincing trend that has occurred is that test-first testing seems to be harder to implement than test-last testing.
+This trend is apparent from the results in Janzen and also from studies~\cite{Aniche:2010, George:2003} which will be discussed in the next section.
 Although these trends seem to be clear, the rest seem to be more muddled.
-Code correctness for example seems to have no clear trend as multiple studies have reached contradictory conclusions to other studies.
-Another hot button topic that fails to show a trend one way or the other is the amount of time spent simply developing code as this again seems to have results across the board.
-From these trends we can consider that test first testing overall probably takes more time to write, increases code coverage, and is harder to implement than test-last testing and that other advantages or detriments may exist but they are unclear in current research.
+Code correctness, for example, seems to have only a vague trend that agrees that test-first methods do not have worse code quality then test-last methods.
+The debate about whether or not test-first testing produces better code correctness though has yet to be resolved due to conflicting data.
+Another muddled trend is that of total development time.
+Currently, the only thing that can be concluded on this topic is that test-first likely takes at least as much time as test-last methods to implement though it is unclear if test-first take significantly longer then test-last methods.
 
-\section{Test Driven Development}
-\mycomment{newly added section, still rough at best}
-
-Test-first testing in most of these studies are being done in various forms of TDD.
-Because of this, the data for test-first testing has the potential to reflect traits of TDD which are independent of any traits in test-first testing.
-Although all the results have a potential to contain this issue one particular result seems to suggest it may be a trait of TDD, not test-first testing.
-This trait is that Test-first testing is more difficult then test last testing.
+\section{Difficulty and TDD}
+Due to TDD's popularity, almost all test first testing studies are done using TDD.
+Because of this, the data for test-first testing has the potential to reflect an attribute of TDD that is not an attribute of test-first testing.
+Although all the results found in the previous section have the potential to contain this issue, one particular result seems to suggest it may be a attribute of TDD and not test-first testing.
+This attribute is the difficulty of test-first testing.
 
 \subsection{Studies noting TDD difficulty}
-Hammond et.
-al states in his research summary paper on TDD that: "TDD remains deceptively simple to describe but deeply challenging to put into practice effectively" `\cite{Hammond:2012}.
-Hammond reaches this conclusion with the use of multiple studies that show implecations that TDD is complex.
-Some of the studies used to draw this conclusion are summarized below.
+In 2012 Hammond et al did a summary paper on the current state of TDD.
+In this paper Hammond et al states that: ``TDD remains deceptively simple to describe but deeply challenging to put into practice effectively"~\cite{Hammond:2012}.
+Hammond et al defended their claim by summarizing several studies briefly that showed  the complexity of TDD.
+In this section we will take a more in depth look at some of the studies Hammond et al used to draw their conclusion.
+We will then in the next section discuss why these studies prove that the complexity is caused by the design of TDD and not test-first testing.
 
-In the study by Janzen 2007, there was an interesting relation between two different measurement aspects of the opinion study.
-It was noted by Janzen that in every experiment, slightly less people said they would be more likely to implement TDD than people who said that TDD was a superior method to test-last methods.
-This means that more people thought that TDD was the better method to implement than  people who said they would actually implement TDD.
-Janzen attributed this phenomena as a testament to the difficulty of TDD as he found in the comments section multiple survey participants mentioned that they felt that TDD was too difficult or too different from what they normally do.
+One of the studies that Hammond et al summarizes is the opinion study by Janzen.
+In particular Hammond focuses on a quote from Janzens study: ``A few of the participants of the studies found TDD either too difficult or too different from what they normally do''.
+This quote refers to an interesting relation found between the choice and best approach category.
+It was noted by Janzen that in every experiment, there were more people who thought test-first was the best approach than people who would choose to implement test-first testing.
+When Janzen looked into why this phenomena had occurred he found multiple statements in the comment section that claimed the difficulties of TDD and attributed the phenomena to TDD difficulty.
 
-In a study done by George and Williams,  \mycomment{cite this article!} multiple programming professionals from John Deere, RoleModel Software, and Ericsson participated in an experiment comparing TDD to waterfall development.
+Another study that Hammond summarizes is a experiment survey by George and Williams~\cite{George:2003}.
+In the study, multiple programming professionals from John Deere, RoleModel Software, and Ericsson participated in an experiment comparing TDD to waterfall development.
 In this study a 9 question survey was given out asking programmers what they thought about TDD and what was difficult with TDD.
@@
 
-In an online survey done by Aniche and Gerosa ~\cite{Aniche:2010}, 218 TDD programmers of differing skill levels were surveyed about their TDD practices and mistakes.
-In the survey it was found that TDD was not easy to follow as about 25\% of the programmers admitted to frequently or always making mistakes in following the traditional steps of TDD.
+The last study Hammond summarizes in this portion of the paper is an online survey done by Aniche and Gerosa~\cite{Aniche:2010}.
+In the survey, 218 TDD programmers of differing skill levels were surveyed about their TDD practices and mistakes.
+In the survey, it was found that TDD was not easy to follow as about 25\% of the programmers admitted to frequently or always making mistakes in following the traditional steps of TDD.
 Two examples of these mistakes include: forgetting to clean up their code after a test passes and writing tests that are too complex for effective TDD.
 
-\subsection{Reasons for TDD Difficulty}
+\subsection{Discussion}
 The data above shows fairly strong evidence that TDD is at least somewhat difficult to implement but none of the data above shows that test-first testing is the problem.
@@
 In the second study, George and Williams reveal that 56\% of the participants found the TDD mindset hard to adapt while only 23\% of the participants found testing before designing as a problem, suggesting that something other then test-first testing caused difficulty for at least 33\% of the participants.
-Whereas the third study shows that programmers have some trouble implementing the steps of TDD, which has nothing to do with testing before designing.
-Since difficulty in using TDD does not seem to be directly linked with test-first testing, that suggests that there is different issue causing TDD to be difficult to implement.
+The third study showed that programmers had trouble implementing the steps of TDD, which has nothing to do with testing before designing.
+With these studies in mind, difficulty in using TDD does not seem to be directly linked with test-first testing.
+This suggests that a different issue is causing TDD to be difficult to implement.
 
-One potential issue with TDD that is brought up by many developers is that it fails to explain the best way to implement tests to test your codes wanted functionality.
-This is best represented in a quote from Dan North's article \mycomment{cite Introducing BDD here}: "While using and teaching agile practices like test-driven development (TDD) on projects in different environments, I kept coming across the same confusion and misunderstandings.
-Programmers wanted to know where to start, what to test and what not to test, how much to test in one go, what to call their tests, and how to understand why a test fails."  This issue as shown in Dan North's article may reveal why TDD is so difficult.
-If the issue for why TDD is so difficult is similar to something like the issue explained above then perhaps they may be a better implementation other than TDD that exists.
+One potential issue with TDD that is brought up by many developers is that it fails to explain how to write good test-first tests.
+This is best represented in a quote from Dan North's article~\cite{North:xxx}: ``While using and teaching agile practices like test-driven development (TDD) on projects in different environments, I kept coming across the same confusions and misunderstandings.
+Programmers wanted to know where to start, what to test and what not to test, how much to test in one go, what to call their tests, and how to understand why a test fails."  In this quote Dan North expresses his and other coders frustration with TDD and its failure to specify how and what tests should be written.
+This problem is also mentioned in articles~\cite{Kettunen:2010, Seoken:2012}.
+Currently their is no empirical evidence that this could be the cause of the TDD difficulty, but it is currently suspect by knowledgeable experts in the field.
 
-\section{evolutions to TDD}
-\mycomment{this section is currently a work in progress}
+\section{Evolutions of TDD}
 
-As can been seen in the above section there is an issue where TDD can be difficult to implement correctly.
-Due to this issue, new methods are starting to appear in the agile community that contain some of TDD's main tenets but have shifted enough away from TDD at some fundamental level that they are starting to receive their own names and classifications.
-Although there are many of TDD spin-offs, for this paper we chose to focus on two more popular spin offs, acceptance test driven development and behavior driven development.
-http://start.fedoraproject.org/
-\mycomment{add lack of research disclaimer}
+New test-first methods are starting to appear in the agile community that contain some of TDD's main tenets but are different enough from TDD that they are starting to receive their own names and classifications.
+These new methods are starting to appear because many developers in one or another are not satisfied with the current implementation of TDD.
+In this subsection we will discuss one of these popular TDD spin-offs, behavior driven development, which was developed to combat the problem noted in the previous paragraph.
 
-\subsection{Acceptance Test Driven Development}
+%\subsection{Acceptance Test Driven Development}
 
-Perhaps the closest spin off to TDD, acceptance test driven development (ATDD), follows many of the ideas and expectations that regular TDD does, like writing tests first and using tests to define the development code.
-Unlike TDD, though, ATDD believes that in order to produce higher quality code, reduce coding confusion, and promote better customer satisfaction, customers should write or define "acceptance" tests.
+%Perhaps the closest spin off to TDD, acceptance test driven development (ATDD), follows many of the ideas and expectations that regular TDD does, like writing tests first and using tests to define the development code.
+Unlike TDD, though, ATDD believes that in order to produce higher quality code and reduce code, customers should write or define "acceptance" tests.
 These acceptance tests are tests that are required to pass before the customer will accept the product and usually define the core functionality of the product being developed.
@@
 
-The shift from having the developer write all the tests to the customer having to write some of tests is ideally considered a win win.
+%The shift from having the developer write all the tests to the customer having to write some of tests is ideally considered a win win.
 The shift of the workload of determining main functionality from the developer to the customer significantly reduces the difficulty of the TDD practice to the developer as they are no longer concerned about writing and determining all the functionality for their code.
@@
 
-\subsection{Behavior Driven Development}
-\mycomment{somewhat stubbed section atm}
-Behavior Driven Development, or BDD, is a new spin off of TDD that focuses on how to correctly implement the fundamental usefulness of TDD.
-The main difference between BDD and TDD is that TDD focuses on "testing the code" while BDD focuses more on defining a codes behavior.
-In order to do this, BDD focuses on defining a codes wanted functionality in plain English and then converts that plain English into tests using various programming tools and methods \mycomment{cite dan North and hammond article here}.
-This differs from TDD, because in TDD, their is no focus on defining code functionality in terms of plain English or even focusing specifically on code functionality.
-
-The argued pros for BDD is that it can bypass many of the difficulties of TDD by allowing us to define the purpose of our code in more natural terms.
-This style removes much of the confusion about what is important to test in TDD and the other pro is that now it is easier to communicate goals and functionality of the code to non computer science majors (IE Managers, Customers, Sales Reps, etc.) which allows for higher customer satisfaction.
+Behavior Driven Development, or BDD for short, was created by Dan North as a substitution for TDD that helps the user determine what to test and how to test it.
+The main difference between BDD and TDD is that TDD focuses on testing the code itself while BDD focuses more on testing a codes intended behavior.
+In order to do this, BDD users start by considering their codes intended functionality.
+They then sentences that use their native language to define what they want their code to achieve.
+These sentences are then converted from the programmers native language to tests in a standard testing language using various programming tools and methods that have been developed specifically for BDD \cite{Soeken:2012, North:xxx, Hammond:2012}.
+This differs from TDD because the focus is no longer on testing new features but on defining what these new features should do.
+This shift in focus and use of native language to define problems, according the Dan North, is what makes BDD simpler to implement.
+\cite{North:xxx}
 
 \section{conclusion}
-\mycomment{this section is currently a work in progress}
-\subsection{Conclusion intro}
+\subsection{Summary}
 
-In this paper we talked about two common test methods: test-first testing, often found in waterfall development models and test-last testing, often found in agile development models.
-In section two we defined these test methods as well as provided background information to the reader.
-In section three we explored recent research into the comparison of the two different testing methods and explored potential research problems with the current data.
-In particular we found that data was contradictory but we still were able to tentatively conclude that test-first testing increases test coverage, increases time taken to test, and is harder to implement than test-last testing.
-In this section we also suggested two potential reasons for the contradictory data: different participant experience levels and poor documentation/implementation of specific test first methods.
-In section four we explored different implementations of test first testing.
+In this paper we talked about and compared two common test methods: test-first testing and test-last testing.
+Within this discussion we reviewed current research in the field in an attempt to make conclusions about the advantages and disadvantages of test-first and test-last methods.
+In this part of the paper we were somewhat able to conclude that test-first testing increases code coverage, is more difficult to implement, has at least as much code correctness, and takes at least as long to develop as test-last testing.
+These conclusions must be considered weak though as we noted that studies were very contradictory to one another due to various factors including method difficulty and poor study design/documentation.
+We also discussed how there seems to be evidence that being harder to implement is not a problem with test-first testing but with TDD, the way test first testing is usually implemented.
+We ended by exploring a new test-first method that had the potential to fix this issue with TDD.
 
-\subsection{Discussi or BDDon of Implications of Section 3}
-\mycomment{stubbed for now}
-
-\subsection{Discussion of Implications of Section 4}
-\mycomment{stubbed for now}
-\subsection{Suggestion for Further Research}
-For people who are looking to advance the field of TDD it would be potentially very interesting to see how the new TDD spin-offs such as Behavior Driven Development and Agile Specification-Driven Development compare to that of a well documented TDD process like the one suggested for Extreme Programming.
-Another useful study that could be implemented would be to compare the results of studies thing that could also help advance the field is to do research comparing the effectiveness of waterfall testing and TDD testing between groups with notably different experience levels as some research tentatively points that this may be a relevant variable.
-
-\mycomment{This paragraph is a rough idea on the further research topic}
-
-\mycomment{conclusion brain storm}
-
--TDD increases code test coverage
-
--TDD perhaps takes more skill to preform
-
--TDD may have more effective and less effective practices
-
--More study is needed.
-
--Better study methods should be implimented
+\subsection{Suggested Further Research}
+For people who are looking to increase clarity in the debate between test-first and test-last testing there are a few studies that would be quite useful in its current state.
+One helpful study that could be implemented is a summary study of TDD and test-first testing articles which are divided up by participant coding expertise to see if participants with different programing experience achieve differing results.
+Another useful study that could be implemented would be to compare the new BDD method against TDD to see if participants find BDD simpler to implement than TDD.
+Also any new research on the debate that has a well documented TDD method and has some way to confirm that participants preformed this method would be useful in helping to solve the debate by giving it more creditable studies to work with.
 

### 2013-11-27 Paper_Draft.tex
Final Paper, barring potential touch ups Sunday.

@@
 
-\title{An Overview of the Current State of Test-First vs. Test-Last Debate}
+\title{An Overview of the Current State of the Test-First vs. Test-Last Debate}
 
@@
 
-Although there are many testing methods that  exist, two particular testing methods are currently very popular: test-last testing and test-first testing.
+Although there are many testing methods that  exist, they can be roughly classified into two categories: test-last testing and test-first testing.
 Test-last testing, used mostly in process oriented or waterfall development, is a testing method where testing is done after software is written to ensure that the software is working as intended.
@@
 The paper is divided into four sections.
-In Section II we discuss what software testing is and current software development models with their supporting testing methods: test-first and test-last testing.
-In Section III we will provide an analysis of the data explaining the potential advantages and disadvantages of test-first testing compared to test-last testing.
-In Section IV we discuss issues of using test driven development to implement test-first testing and will go over a new test-first methodology called behavior driven development.
-In Section V we will provide conclusions and suggestions for further research in the field.
+In Section 2 we discuss what software testing is and current software development models with their supporting testing methods: test-first and test-last testing.
+In Section 3 we will provide an analysis of the data explaining the potential advantages and disadvantages of test-first testing compared to test-last testing.
+In Section 4 we discuss issues of using test driven development to implement test-first testing and will go over a new test-first methodology called behavior driven development.
+In Section 5 we will provide conclusions and suggestions for further research in the field.
 
@@
 Code coverage refers to the percentage of lines of production code that are executed when a set of tests are run.
-For example, if I write a set of tests that has 70\% test coverage, it means that 70\% of the lines of production code were executed when the tests were run.
+For example, if one writes a set of tests that has 70\% test coverage, it means that 70\% of the lines of production code were executed when the tests were run.
 Total development time refers to how long it takes to finish a development project.
@@
 In response to these critiques of the waterfall model, a new model for developing software emerged, the agile development model.
-This new development model, based on the tenets of the agile manifesto ~\cite{agile:xxx}, promoted the idea that all actions of development should not occur within an ordered sequence of phases, but instead a series of time-boxed iterations where, in each iteration, developers set requirements, design a product, make the product, and test the product based on feedback from the previous iteration.
+This new development model, based on the tenets of the agile manifesto ~\cite{agile:xxx}, promoted the idea that all actions of development should not occur within an ordered sequence of phases, but instead a series of time-boxed iterations where, in each iteration, developers set requirements, design, make , and test a subset of the end product's core functionality based on feedback from the previous iteration.
 The goal of each iteration is to produce a demonstrable sub-product to show a customer and to receive feedback on that sub-product.
+This process is complete when the sub-product meets all the demands of the customer and becomes the end product.
 Some current development practices that are considered agile are Extreme Programming and Scrum.
@@
 We will first review three main studies.
-Next, we will summarize the results of our studies in terms of our three main empirical measurements: code coverage, code correctness, and total development time.
-Afterwards we will discuss the potential issues with summarizing the data given in the research articles and then we will talk about potential useful comparisons we can draw from the data.
+After that we will discuss the potential issues with summarizing the data given in the research articles and then draw potential useful comparisons from the data.
 
@@
 This subsection contains summaries of important studies that will be used later in the paper to draw out important conclusions and comparisons between test-first and test-last testing.
-During this section we will make multiple references to test last development methods which we will refer to as TLD.
+During this section we will make multiple references to test-last development methods which we will refer to as TLD.
 
@@
 First, is a 2010 review by Kollanus \cite{Kollanus:2010}.
-In this review, Kollanus,  reviewed forty different experiments in scientific journals, magazines, and conference proceedings that provided empirical evidence comparing TDD  to TLD.
+In this review, Kollanus reviewed forty different experiments in scientific journals, magazines, and conference proceedings that provided empirical evidence comparing TDD  to TLD.
 Each study was assigned to one of three categories: Controlled experiments, case studies, and others.
@@
 In this review, Kollanus focuses on three different code quality measurements: external code quality, productivity, and internal code quality.
-External code quality, as defined by Kollanus, refers to how many errors are found in participants code.
+External code quality, as defined by Kollanus, refers to how many errors are found in the resulting code.
 This definition is the same as the definition for code correctness.
@@
 
-In terms of productivity, Kollanas suggests that TDD \textit{may} be less productive.
+In terms of productivity, Kollanus suggests that TDD \textit{may} be less productive.
 This conclusion was drawn from 23 studies where 11 of the studies claimed that TDD decreases productivity, 7 of the studies say that there was no difference in productivity, and 5 studies that say there was an increase of productivity.
@@
 This study used 39 third-year computer science students knowledgeable in testing techniques.
-Each student took part in  two 100 minute test-first modules and were then asked to complete coding challenges over two sessions.
+Each student took part in  two 100 minute test-first training modules and were then asked to complete coding challenges over two sessions.
 In the first session, half the students used test-first methods while the other half used test-last methods.
@@
 
-Code coverage and total development time were measured as mentioned in Section II.
+Code coverage and total development time were measured as mentioned in Section 2.
 Code correctness was measured by running a set acceptance tests designed by the experimenter against a participant's code.
@@
 On total development time, it was found that test-first code took 12\% longer to write then test-last code.
-This result was found to be significantly significant which lead to the conclusion that test-first methods took longer to implement then test-last methods.
+This result was found to be statistically significant which lead to the conclusion that test-first methods took longer to implement then test-last methods.
 In terms of code correctness, the only difference found was that the test-last code had one more correct implementation of code then the test-first code.
@@
 
-Last, is a 2007 opinion study by Janzen \cite{Janzen:2007}.
+In 2007 an opinion study was conducted by Janzen \cite{Janzen:2007}.
 In this opinion study, Janzen polled participants from six experiments to determine whether they preferred testing with test-first methods or test-last methods.
@@
 
-The results were divided between a beginner group (students within their first few code classes) and a mature group (students nearing undergraduate graduation and professionals).
-The results for the beginner group can be seen in figure 1 and the results for the mature group can be seen in figure 2.
+The results were divided between a beginner group (students within their first few programming classes) and a mature group (students nearing undergraduate graduation and professionals).
+The results for the beginner group can be seen in Figure 1 and the results for the mature group can be seen in Figure 2.
 Each graph shows the distribution of all of the participants answers to the six questions where each question is referenced by its bold keyword.
-With the data given in the figures, Janzen was able to conclude that beginner coding students have a preference toward test-last testing while more mature coding students and industry professionals show no significant preference toward either testing method.
+With the data given in the figures on the choice question, Janzen was able to conclude that beginner programming students have a statistically significant preference toward using test-last testing while more mature programming students and industry professionals showed a preference toward using test-first testing though preference was not found to be statistically significant.
 
@@
 This is a concern because most study data comes from two sources with very different skill levels: industry professionals and college students.
-In the study by Janzen et al, figures 1 and 2 show very different results between beginning and mature programmers.
-In particular, beginner programmers did not report a very positive experience with test-first testing compared mature programmers.
+In the study by Janzen, figures 1 and 2 show very different results between beginning and mature programmers.
+In particular, beginner programmers did not report an overall beneficial experience with test-first testing compared mature programmers.
 This suggests that beginning coders may not be getting all the benefits from test-first methods that the more mature coders received.
@@
 
-The other problem that has plagued summary papers of test first testing is the lack documentation  and conformance of TDD methods in studies.
+The other problem that has plagued summary papers of test-first testing is the lack documentation  and conformance of TDD methods in studies.
 Kollanus, in her review paper, acknowledges a frustrating lack of information in many studies on how TDD was implemented.
@@
 Due to current issues in contradictory data it is hard to make solid conclusions about the advantages and disadvantages of using test-first testing instead of test-last testing.
-That being said current trends exist within the research that allow us to guess some of the traits of test-first methods compared to test-last methods.
+That being said, current trends exist within the research that allow us to guess some of the traits of test-first methods compared to test-last methods.
 One result which was very clear in the research was that of code coverage.
@@
 Although these trends seem to be clear, the rest seem to be more muddled.
-Code correctness, for example, seems to have only a vague trend that agrees that test-first methods do not have worse code quality then test-last methods.
+Code correctness, for example, seems to have only a vague trend that agrees that test-first methods do not have worse code quality than test-last methods.
 The debate about whether or not test-first testing produces better code correctness though has yet to be resolved due to conflicting data.
 Another muddled trend is that of total development time.
-Currently, the only thing that can be concluded on this topic is that test-first likely takes at least as much time as test-last methods to implement though it is unclear if test-first take significantly longer then test-last methods.
+Currently, the only thing that can be concluded on this topic is that test-first likely takes at least as much time as test-last methods to implement though it is unclear if test-first methods take significantly longer than test-last ones.
 
 \section{Difficulty and TDD}
-Due to TDD's popularity, almost all test first testing studies are done using TDD.
+Due to TDD's popularity, almost all test first testing studies claim to have been done using TDD.
 Because of this, the data for test-first testing has the potential to reflect an attribute of TDD that is not an attribute of test-first testing.
@@
 In 2012 Hammond et al did a summary paper on the current state of TDD.
-In this paper Hammond et al states that: ``TDD remains deceptively simple to describe but deeply challenging to put into practice effectively"~\cite{Hammond:2012}.
+In this paper Hammond et al states that ``TDD remains deceptively simple to describe but deeply challenging to put into practice effectively"~\cite{Hammond:2012}.
 Hammond et al defended their claim by summarizing several studies briefly that showed  the complexity of TDD.
-In this section we will take a more in depth look at some of the studies Hammond et al used to draw their conclusion.
-We will then in the next section discuss why these studies prove that the complexity is caused by the design of TDD and not test-first testing.
+In this section we will take a more in-depth look at some of the studies Hammond et al used to draw their conclusion.
+We will then discuss in the next section why these studies prove that the complexity is caused by the design of TDD and not test-first testing.
 
@@
 It was noted by Janzen that in every experiment, there were more people who thought test-first was the best approach than people who would choose to implement test-first testing.
-When Janzen looked into why this phenomena had occurred he found multiple statements in the comment section that claimed the difficulties of TDD and attributed the phenomena to TDD difficulty.
+When Janzen looked into why this phenomena had occurred he found multiple statements in the comment section that claimed the difficulties of TDD.
 
-Another study that Hammond summarizes is a experiment survey by George and Williams~\cite{George:2003}.
+Another study that Hammond summarizes is an experiment survey by George and Williams~\cite{George:2003}.
 In the study, multiple programming professionals from John Deere, RoleModel Software, and Ericsson participated in an experiment comparing TDD to waterfall development.
 In this study a 9 question survey was given out asking programmers what they thought about TDD and what was difficult with TDD.
-In this survey 56\% of the professionals noted that they had difficulty adapting to the TDD mindset when participating the study.
-In addition 23\% of the participants noted that they felt that the lack of upfront design in TDD was more of a hindrance then a help.
+In this survey 56\% of the professionals noted that they had difficulty adapting to the TDD mindset when participating in the study.
+In addition 23\% of the participants noted that they felt that the lack of upfront design in TDD was more of a hindrance than a help.
 
@@
 With these studies in mind, difficulty in using TDD does not seem to be directly linked with test-first testing.
-This suggests that a different issue is causing TDD to be difficult to implement.
 
@@
 Programmers wanted to know where to start, what to test and what not to test, how much to test in one go, what to call their tests, and how to understand why a test fails."  In this quote Dan North expresses his and other coders frustration with TDD and its failure to specify how and what tests should be written.
-This problem is also mentioned in articles~\cite{Kettunen:2010, Seoken:2012}.
-Currently their is no empirical evidence that this could be the cause of the TDD difficulty, but it is currently suspect by knowledgeable experts in the field.
+This problem is also mentioned in articles~\cite{Kettunen:2010, Soeken:2012}.
+Currently there is no empirical evidence that this could be the cause of the TDD difficulty, but it is currently suspect by knowledgeable experts in the field~\cite{Soeken:2012, Kettunen:2010, North:xxx}.
 
-\section{Evolutions of TDD}
+\section{Evolutions of TDD: BDD}
 
 New test-first methods are starting to appear in the agile community that contain some of TDD's main tenets but are different enough from TDD that they are starting to receive their own names and classifications.
-These new methods are starting to appear because many developers in one or another are not satisfied with the current implementation of TDD.
+These new methods are starting to appear because many developers are not satisfied with the current implementation of TDD.
 In this subsection we will discuss one of these popular TDD spin-offs, behavior driven development, which was developed to combat the problem noted in the previous paragraph.
@@
 Behavior Driven Development, or BDD for short, was created by Dan North as a substitution for TDD that helps the user determine what to test and how to test it.
-The main difference between BDD and TDD is that TDD focuses on testing the code itself while BDD focuses more on testing a codes intended behavior.
-In order to do this, BDD users start by considering their codes intended functionality.
-They then sentences that use their native language to define what they want their code to achieve.
+The main difference between BDD and TDD is that TDD focuses on testing the code itself while BDD focuses more on testing a code's intended behavior.
+In order to do this, BDD users start by considering their code's intended functionality.
+They then write sentences that use their native language to define what they want their code to achieve.
 These sentences are then converted from the programmers native language to tests in a standard testing language using various programming tools and methods that have been developed specifically for BDD \cite{Soeken:2012, North:xxx, Hammond:2012}.
 This differs from TDD because the focus is no longer on testing new features but on defining what these new features should do.
-This shift in focus and use of native language to define problems, according the Dan North, is what makes BDD simpler to implement.
-\cite{North:xxx}
+This shift in focus and use of native language to define problems, according the Dan North, is what makes BDD simpler to implement \cite{North:xxx}.
+However, there have not been enough studies done to check this hypothesis.
 
@@
 These conclusions must be considered weak though as we noted that studies were very contradictory to one another due to various factors including method difficulty and poor study design/documentation.
-We also discussed how there seems to be evidence that being harder to implement is not a problem with test-first testing but with TDD, the way test first testing is usually implemented.
-We ended by exploring a new test-first method that had the potential to fix this issue with TDD.
+We also discussed how there seems to be evidence that being harder to implement is not a problem with test-first testing but with TDD, the way test-first testing is usually implemented.
+We concluded by exploring a new test-first method, BDD, that had the potential to fix this issue with TDD.
 
@@
 Another useful study that could be implemented would be to compare the new BDD method against TDD to see if participants find BDD simpler to implement than TDD.
-Also any new research on the debate that has a well documented TDD method and has some way to confirm that participants preformed this method would be useful in helping to solve the debate by giving it more creditable studies to work with.
+Also any new research on the debate that has a well documented TDD method and has some way to confirm that participants followed this method would increase the amount of credible data on TDD.
 

### 2013-11-27 Paper_Draft.tex
Corrected a quick issue in the paper.

@@
 In 2007 an opinion study was conducted by Janzen \cite{Janzen:2007}.
-In this opinion study, Janzen polled participants from six experiments to determine whether they preferred testing with test-first methods or test-last methods.
+In this opinion study, Janzen polled participants from six TDD experiments to determine whether they preferred testing with test-first methods or test-last methods.
 Five of these experiments were conducted on students at the University of Kansas while one of the experiments was conducted on professional programmers from a Fortune 500 company.
@@
 One of the studies that Hammond et al summarizes is the opinion study by Janzen.
-In particular Hammond focuses on a quote from Janzens study: ``A few of the participants of the studies found TDD either too difficult or too different from what they normally do''.
+In particular Hammond focuses on a quote from Janzens study: ``They found that while some programmers saw the benefits of the test-first approach, several of the programmers had the perception that it was too difficult or too different from what they normally do''.
 This quote refers to an interesting relation found between the choice and best approach category.
 It was noted by Janzen that in every experiment, there were more people who thought test-first was the best approach than people who would choose to implement test-first testing.
-When Janzen looked into why this phenomena had occurred he found multiple statements in the comment section that claimed the difficulties of TDD.
+When Janzen looked into why this phenomena had occurred he found multiple statements in the comment section that claimed the difficulties of implementing test-first methods (which in this study was TDD).
 
@@
 The data above shows fairly strong evidence that TDD is at least somewhat difficult to implement but none of the data above shows that test-first testing is the problem.
-In the first study, Janzen reveals that TDD was difficult for some people to implement.
+In the first study, Janzen reveals that test-first methods, in the form of TDD, were difficult for some people to implement.
 In the second study, George and Williams reveal that 56\% of the participants found the TDD mindset hard to adapt while only 23\% of the participants found testing before designing as a problem, suggesting that something other then test-first testing caused difficulty for at least 33\% of the participants.
